### Part 4: Generating comparison dataset at the pair level taking the publication into account

```{r}
prepubpaircomp_table_1 <- unique(merge(imex_human_sel,reactome_pairs_pmids,by=c("pair_id","pmid"),all=T))
prepubpaircomp_table_2 <- unique(merge(prepubpaircomp_table_1,tm_pairs_pmids_sel,by=c("pair_id","pmid"),all=T))
pubpaircomp_table <- unique(merge(prepubpaircomp_table_2,iid_pred_pairs,by=c("pair_id"),all=T))

# I clean and replace all NAs if present.

pubpaircomp_table_final <- pubpaircomp_table
pubpaircomp_table_final[is.na(pubpaircomp_table_final <- pubpaircomp_table)] <- 0
write.table(pubpaircomp_table_final,"./results/pubpaircomp_table_final.txt",col.names=T,row.names=F,sep="\t",quote=F)
```

The comparison set gives a total number of `r nrow(pubpaircomp_table_final)` protein association pairs, of which `r nrow(pubpaircomp_table_final[pubpaircomp_table_final$imex==0,])` (`r round(nrow(pubpaircomp_table_final[pubpaircomp_table_final$imex==0,])*100/nrow(pubpaircomp_table_final),2)`%) are not curated in IMEx. In all these pairs the publication from which they were derived was also matched, so the overlaps and numbers differ from my previous comparisons. 

I produce a plot with the summary of the overlap between the different datasets evaluated. 

##### Figure 5: Comparison between different protein association datasets at the pair+publication level
```{r echo=FALSE, fig.width=10,fig.height=6}
upset(pubpaircomp_table_final, 
      nsets = 4, 
      point.size = 6, 
      name.size = 12, 
      line.size = 2, 
      mainbar.y.label = "Common protein pairs", 
      sets.x.label = "Nr of protein pairs in dataset", 
      order.by="freq",
      decreasing=FALSE,
      queries = list(
              list(query = intersects, params = list("reactome","iid_pred","tm_epmc"), color = "blue", active = T),
              list(query = intersects, params = list("iid_pred","tm_epmc"), color= "cornflowerblue",active = T),
              list(query = intersects, params = list("reactome","iid_pred"), color= "cornflowerblue",active = T),
              list(query = intersects, params = list("reactome","tm_epmc"), color= "cornflowerblue",active = T),
              list(query = intersects, params = list("imex","iid_pred","reactome","tm_epmc"), color= "darkorange2",active = T),
              list(query = intersects, params = list("imex","reactome","tm_epmc"), color= "orange1",active = T),
              list(query = intersects, params = list("imex","iid_pred","tm_epmc"), color= "orange1",active = T),
              list(query = intersects, params = list("imex","iid_pred","reactome"), color= "orange1",active = T),
              list(query = intersects, params = list("imex","iid_pred"), color= "lightgoldenrod2",active = T),
              list(query = intersects, params = list("imex","reactome"), color= "lightgoldenrod2",active = T),
              list(query = intersects, params = list("imex","tm_epmc"), color= "lightgoldenrod2",active = T),
              list(query = intersects, params = list("imex"), color= "gray70",active = T),
              list(query = intersects, params = list("reactome"), color= "gray70",active = T),
              list(query = intersects, params = list("iid_pred"), color= "gray70",active = T),
              list(query = intersects, params = list("tm_epmc"), color= "gray70",active = T)))
```

And again, a sub-plot showing exclusively the intersections, for better comparison.  

##### Figure 6: Intersection between different protein association datasets at the pair+publication level
```{r echo=FALSE,fig.width=10,fig.height=6}
upset(pubpaircomp_table_final, 
      nsets = 4, 
      point.size = 6, 
      name.size = 12, 
      line.size = 2, 
      mainbar.y.label = "Common protein pairs", 
      sets.x.label = "Nr of protein pairs in dataset",
      order.by="freq",
      decreasing=FALSE,
      queries = list(
              list(query = intersects, params = list("reactome","iid_pred","tm_epmc"), color = "blue", active = T),
              list(query = intersects, params = list("iid_pred","tm_epmc"), color= "cornflowerblue",active = T),
              list(query = intersects, params = list("reactome","iid_pred"), color= "cornflowerblue",active = T),
              list(query = intersects, params = list("reactome","tm_epmc"), color= "cornflowerblue",active = T),
              list(query = intersects, params = list("imex","iid_pred","reactome","tm_epmc"), color= "darkorange2",active = T),
              list(query = intersects, params = list("imex","reactome","tm_epmc"), color= "orange1",active = T),
              list(query = intersects, params = list("imex","iid_pred","tm_epmc"), color= "orange1",active = T),
              list(query = intersects, params = list("imex","iid_pred","reactome"), color= "orange1",active = T),
              list(query = intersects, params = list("imex","iid_pred"), color= "lightgoldenrod2",active = T),
              list(query = intersects, params = list("imex","reactome"), color= "lightgoldenrod2",active = T),
              list(query = intersects, params = list("imex","tm_epmc"), color= "lightgoldenrod2",active = T)),
      intersections = list(
              list("reactome","iid_pred","tm_epmc"),
              list("iid_pred","tm_epmc"),
              list("reactome","iid_pred"),
              list("reactome","tm_epmc"),
              list("imex","iid_pred","reactome","tm_epmc"),
              list("imex","reactome","tm_epmc"),
              list("imex","iid_pred","tm_epmc"),
              list("imex","iid_pred","reactome"),
              list("imex","iid_pred"),
              list("imex","reactome"),
              list("imex","tm_epmc")))
```


Checking for those pmids that have been curated in IMEx but for which the pairs do not match with the other resources. 
```{r}


pubpair_poterrors <- unique(merge(pubpaircomp_table,imex_pmids,all=F))
pubpair_poterrors <- unique(select(pubpair_poterrors,pmid))
pubpair_poterrors$pot_pred_error <- 1

pubpaircomp_table_errors <- unique(merge(pubpaircomp_table,pubpair_poterrors,all=T))

```
