---
title: "dsp_comparison_groups_enrichment_pres"
author: "Vitalii Kleshchevnikov"
date: "4/12/2017"
output:
  ioslides_presentation: default
  slidy_presentation: default
---

```{r set-options, echo=FALSE}
options(width = 80, defaultWidth = 1000, defaultHeight = 1200, echo = F)
#runtime: shiny
```

Estimating the size of the uncurated interactome
========================================================

### Synopsis

After producing tidy datasets comparing different resources to IMEx data, we put together the data and compare the overlap. 

IMPORTANT: This set of scripts assume the different referenced datasets have been freshly updated. If you need to update results of this set, please re-run the corresponding source as well. 

# Loading and processing data

## Required libraries
```{r libraries,message=FALSE,warning=FALSE}
# install packages if some are not already installed
packages = c("plyr","dplyr","data.table","UpSetR", "splitstackshape", "ggplot2", "shiny", "VennDiagram")
if(mean(packages %in% installed.packages()) != 1){
        install.packages(packages[!packages %in% installed.packages()])
}
suppressPackageStartupMessages({
library(plyr)
library(dplyr)
library(data.table)
library(UpSetR)
library(splitstackshape)
library(ggplot2)
library(shiny)
library(VennDiagram)
})
```

## Part 1: Load datasets

### IMEx dataset

I select only purely human interactions here (interactions where both proteins are human). 

```{r warning=FALSE,message=FALSE}
imex_full <- fread("../IMEx/results/imex_full.txt", header=T, sep="\t",colClasses="character",data.table = T)
imex_human <- unique(subset(imex_full,taxid_a=="9606" & taxid_b=="9606"))
imex_human$imex <- 1
imex_human_sel <- unique(select(imex_human,pair_id=pair_id_clean,pmid=pubid,imex))
imex_pairs <- unique(select(imex_human,pair_id=pair_id_clean,imex))
imex_pmids <- unique(select(imex_human,pmid=pubid,imex))
```

The dataset contains `r nrow(imex_pairs)` protein interactions recorded in `r nrow(imex_pmids)` publications.

### Reactome data

```{r, fig.height=8, fig.width=8}
setwd("../reactome_interactions/results/")
system("gunzip -k pairs_pmid_reactome.txt.gz")
reactome_pairs_pmids <- fread("pairs_pmid_reactome.txt",header=T,sep="\t",colClasses=c("character","character","numeric"),data.table = T)
system("rm pairs_pmid_reactome.txt")
setwd("../../dsp_comparison/")

reactome_pairs <- unique(select(reactome_pairs_pmids,pair_id,reactome))
reactome_pmids <- unique(select(reactome_pairs_pmids,pmid,reactome))
```

The reactome dataset contains `r nrow(reactome_pairs)` protein associations recorded in `r nrow(reactome_pmids)` publications. 

### Text-mining EPMC data

```{r, fig.height=8, fig.width=8}
setwd("../epmc_text_mining/results/")
system("gunzip -k pairs_pmids_tm.txt.gz")
tm_pairs_pmids <- fread("pairs_pmids_tm.txt",header=T,sep="\t",colClasses=c("character","character","numeric"),data.table = T)
system("rm pairs_pmids_tm.txt")
setwd("../../dsp_comparison/")

tm_pairs_pmids_sel <- unique(select(tm_pairs_pmids,pair_id,pmid,tm_epmc=tm))
tm_pairs <- unique(select(tm_pairs_pmids,pair_id,tm_epmc=tm))
tm_pmids <- unique(select(tm_pairs_pmids,pmid,tm_epmc=tm))
```

The text-mining EPMC dataset contains `r nrow(tm_pairs)` protein associations recorded in `r nrow(tm_pmids)` publications.

### IID predictions data

```{r, fig.height=8, fig.width=8}
iid_pred_pairs <- fread("../iid_predictions/results/pairs_iid_pred.txt",header=T,sep="\t",colClasses=c("character","numeric"),data.table=T)
```

The IID-predictions dataset contains `r nrow(iid_pred_pairs)` protein associations.

### Text-mining EVEX data

```{r, fig.height=8, fig.width=8}
setwd("../EVEX/results/")
EVEX_pairs_pmids <- fread("pairs_pmids_EVEX_shallow.txt",header=T,sep="\t",colClasses=c("character","character","character","character","character","numeric"),data.table = T)
setwd("../../dsp_comparison/")

EVEX_pairs_pmids_sel <- unique(select(EVEX_pairs_pmids, pair_id=pair_id_clean, pmid=pubid,EVEX))
EVEX_pairs <- unique(select(EVEX_pairs_pmids, pair_id=pair_id_clean,EVEX))
EVEX_pmids <- unique(select(EVEX_pairs_pmids, pmid=pubid,EVEX))
```

The text-mining EVEX dataset contains `r nrow(EVEX_pairs)` protein associations recorded in `r nrow(EVEX_pmids)` publications.

### BioGRID data

```{r, fig.height=8, fig.width=8}
setwd("../BioGRID/results/")
BioGRID_pairs_pmids <- fread("pairs_pmids_biogrid.txt",header=T,sep="\t",colClasses=c("character","character","numeric"),data.table = T)
setwd("../../dsp_comparison/")

BioGRID_pairs_pmids_sel <- unique(select(BioGRID_pairs_pmids, pair_id=pair_id_clean, pmid=pubid, BioGRID = biogrid))
BioGRID_pairs <- unique(select(BioGRID_pairs_pmids, pair_id=pair_id_clean,BioGRID = biogrid))
BioGRID_pmids <- unique(select(BioGRID_pairs_pmids, pmid=pubid,BioGRID = biogrid))
```

The BioGRID dataset contains `r nrow(BioGRID_pairs)` protein associations recorded in `r nrow(BioGRID_pmids)` publications.

### GO IPI data (EBI_GOA_nonIntAct)

```{r, fig.height=8, fig.width=8}
setwd("../GO_IPI/results/")
GO_IPI_pairs_pmids <- fread("pairs_pmids_EBI_GOA_nonIntAct.txt",header=T,sep="\t",colClasses=c("character","character","numeric"),data.table = T)
setwd("../../dsp_comparison/")

GO_IPI_pairs_pmids_sel <- unique(select(GO_IPI_pairs_pmids, pair_id=pair_id_clean, pmid=pubid, GO_IPI = EBI_GOA_nonIntAct))
GO_IPI_pairs <- unique(select(GO_IPI_pairs_pmids, pair_id=pair_id_clean,GO_IPI = EBI_GOA_nonIntAct))
GO_IPI_pmids <- unique(select(GO_IPI_pairs_pmids, pmid=pubid,GO_IPI = EBI_GOA_nonIntAct))
```

The GO IPI dataset contains `r nrow(GO_IPI_pairs)` protein associations recorded in `r nrow(GO_IPI_pmids)` publications.

### OmniPath interaction and ptm (post-translational modifications) data 

```{r, fig.height=8, fig.width=8}
setwd("../OmniPath/results/")
OmniPath_interaction_pairs_pmids <- fread("pairs_pmids_OmniPath_interactions_minimal.txt",header=T,sep="\t",colClasses=c("character","character","character","character","character","numeric"),data.table = T)
setwd("../../dsp_comparison/")

setwd("../OmniPath/results/")
OmniPath_ptm_pairs_pmids <- fread("pairs_pmids_OmniPath_ptm_interactions_minimal.txt",header=T,sep="\t",colClasses=c("character","character","character","character","character","numeric"),data.table = T)
setwd("../../dsp_comparison/")

OmniPath_interaction_pairs_pmids_sel <- unique(select(OmniPath_interaction_pairs_pmids, pair_id=pair_id_clean, pmid=pubid, OmniPath_interactions))
OmniPath_interaction_pairs <- unique(select(OmniPath_interaction_pairs_pmids, pair_id=pair_id_clean,OmniPath_interactions))
OmniPath_interaction_pmids <- unique(select(OmniPath_interaction_pairs_pmids, pmid=pubid,OmniPath_interactions))

OmniPath_ptm_pairs_pmids_sel <- unique(select(OmniPath_ptm_pairs_pmids, pair_id=pair_id_clean, pmid=pubid, OmniPath_ptm))
OmniPath_ptm_pairs <- unique(select(OmniPath_ptm_pairs_pmids, pair_id=pair_id_clean,OmniPath_ptm))
OmniPath_ptm_pmids <- unique(select(OmniPath_ptm_pairs_pmids, pmid=pubid,OmniPath_ptm))
```

The OmniPath interaction dataset contains `r nrow(OmniPath_interaction_pairs)` protein associations recorded in `r nrow(OmniPath_interaction_pmids)` publications.

The OmniPath ptm dataset contains `r nrow(OmniPath_ptm_pairs)` protein associations recorded in `r nrow(OmniPath_ptm_pmids)` publications.

```{r omnipath_merge}
setnames(OmniPath_interaction_pairs_pmids, colnames(OmniPath_interaction_pairs_pmids)[6], "OmniPath")
setnames(OmniPath_ptm_pairs_pmids, colnames(OmniPath_ptm_pairs_pmids)[6], "OmniPath")
OmniPath_pairs_pmids = rbind(OmniPath_interaction_pairs_pmids, OmniPath_ptm_pairs_pmids)

OmniPath_pairs_pmids_sel <- unique(select(OmniPath_pairs_pmids, pair_id=pair_id_clean, pmid=pubid, OmniPath))
OmniPath_pairs <- unique(select(OmniPath_pairs_pmids, pair_id=pair_id_clean,OmniPath))
OmniPath_pmids <- unique(select(OmniPath_pairs_pmids, pmid=pubid,OmniPath))
```

The OmniPath merged dataset contains `r nrow(OmniPath_pairs)` protein associations recorded in `r nrow(OmniPath_pmids)` publications.

### STRING: textmining, pathway inference, orthology- and comparative genomics-based predictions

```{r STRING}
# read STRING_textmining table
setwd("../STRING/results/")
STRING_textmining <- fread("pairs_STRING_textmining.txt",header=T,sep="\t",colClasses=c("character","character","character","character","numeric","character","numeric"),data.table = T)
STRING_textmining_pairs <- unique(select(STRING_textmining, pair_id=pair_id_clean,STRING_textmining))

# read STRING_pathway_inference table
STRING_pathway_inference <- fread("pairs_STRING_pathway_inference.txt",header=T,sep="\t",colClasses=c("character","character","character","character","numeric","character","numeric"),data.table = T)
STRING_pathway_inference_pairs <- unique(select(STRING_pathway_inference, pair_id=pair_id_clean,STRING_pathway_inference))

# read STRING_phylo_predictions table
STRING_phylo_predictions <- fread("pairs_STRING_phylo_predictions.txt",header=T,sep="\t",colClasses=c("character","character","character","character","numeric","character","numeric"),data.table = T)
STRING_phylo_predictions_pairs <- unique(select(STRING_phylo_predictions, pair_id=pair_id_clean,STRING_phylo_predictions))

setwd("../../dsp_comparison/")

```

 - STRING textmining dataset: contains `r nrow(STRING_textmining_pairs)` protein associations.

 - STRING pathway-based inference dataset: contains `r nrow(STRING_pathway_inference)` protein associations.

 - STRING predictions dataset: contains `r nrow(STRING_phylo_predictions_pairs)` protein associations.

## Part 2: Generating comparison dataset at the pair level

```{r, fig.height=8, fig.width=8}
# Code below generates (to allow any number and any column names) and evaluates this:
# paste0("all_df <- list(",paste0(grep("_pairs$", ls(), value = T), collapse= ","),")")
# eval(parse(text=paste0("all_df <- list(",paste0(grep("_pairs$", ls(), value = T), collapse= ","),")")))
# results in errors upstream

all_df <- list(imex_pairs,reactome_pairs,tm_pairs,iid_pred_pairs, EVEX_pairs, BioGRID_pairs, GO_IPI_pairs, OmniPath_pairs, STRING_textmining_pairs, STRING_pathway_inference_pairs, STRING_phylo_predictions_pairs)

comp_table <- Reduce(function(...) merge(..., all=TRUE), all_df)

# I clean and replace all NAs if present.

comp_table_final <- comp_table
comp_table_final[is.na(comp_table_final <- comp_table)] <- 0
comp_table_final = unique(comp_table_final)
fwrite(comp_table_final,"./results/comp_table_final.txt",col.names=T,row.names=F,sep="\t",quote=F)
system("gzip ./results/comp_table_final.txt --force")
unlink("./results/comp_table_final.txt")
```

The comparison set gives a total number of `r nrow(comp_table_final)` potentially interacting pairs, of which `r nrow(comp_table_final[comp_table_final$imex==0,])` (`r round(nrow(comp_table_final[comp_table_final$imex==0,])*100/nrow(comp_table_final),2)`%) are not curated in IMEx. 

I produce a plot with the summary of the overlap between the different datasets evaluated. 

```{r visualize_overlaps, fig.height=6, fig.width=8, results='hide'}
groups = colnames(comp_table_final)[2:ncol(comp_table_final)]
# generate expression like this "comp_table_final[, overlaps := paste0(imex,reactome,tm_epmc,iid_pred,EVEX,BioGRID,GO_IPI,OmniPath)]" and evaluate it - this way the code is independent of column names
eval(parse(text=paste0("comp_table_final[, overlaps := paste0(",paste0(groups, collapse= ","),")]")))
eval(parse(text=paste0("comp_table_final[, N_overlaps := sum(",paste0(groups, collapse= ","),"), by = pair_id]")))

comp_table_final[, N_per_overlaps := .N, by = overlaps]
comp_table_final[, N_per_N_overlaps := .N, by = N_overlaps]

N_overlaps = unique(comp_table_final[,.(N_overlaps, N_per_N_overlaps)])
N_overlaps = N_overlaps[order(N_per_N_overlaps, decreasing = T),]
qplot(label = N_overlaps$N_per_N_overlaps, y = N_overlaps$N_per_N_overlaps, x = N_overlaps$N_overlaps,  geom = "text") + scale_y_log10(labels = 10^c(1:7), breaks = 10^c(1:7)) +
        ggtitle("The number of interacting pairs and the number of resources \n the pairs are shared between") + ylab("the number of interacting pairs, log10 scale")+ xlab("the number of resources")+scale_x_continuous(labels = N_overlaps$N_overlaps, breaks = N_overlaps$N_overlaps)

overlaps = unique(comp_table_final[,.(overlaps, N_per_overlaps)])
overlaps = overlaps[order(N_per_overlaps, decreasing = T),]
```