---
title: "PMIDs for proteins which have no interactions in IMEx"
author: "Vitalii Kleshchevnikov"
date: "2 May 2017"
output: 
  html_document: 
    keep_md: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# install packages if some are not already installed
# packages = c("plyr","dplyr","data.table","UpSetR", "splitstackshape", "ggplot2", "shiny", "VennDiagram")
# if(mean(packages %in% installed.packages()) != 1){
#        install.packages(packages[!packages %in% installed.packages()])
#}
suppressPackageStartupMessages({
library(data.table)
library(ggplot2)
library(VennDiagram)
library(UniProt.ws)
})
```

#### This script generates the list of the pubmed publication identifiers associated with proteins which have no interactions in any of the IMEx databases including DIP. IMEx dataset is generated by “IMEx dataset generator” pipeline from DSP. It uses two sources of gene/protein to PMID mapping: NCBI and Uniprot.

### protein to pmids mapping from NCBI

```{r load_IMEx, results='hide'}
imex_full <- fread("../IMEx/results/imex_full.txt", header=T, sep="\t",colClasses="character",data.table = T)
imex_human <- unique(subset(imex_full,taxid_a=="9606" & taxid_b=="9606"))
imex_human$imex <- 1
imex_human_proteins = unique(base::c(imex_human$id_a_clean,imex_human$id_b_clean))
```

#### Load gene to PMID dataset (NCBI)

```{r load_NCBI_gene_to_PMID, results='hide'}
geneID2pubmed_url = "ftp://ftp.ncbi.nih.gov/gene/DATA/gene2pubmed.gz"
geneID2pubmed_filename = paste0("./geneID2pubmed_release_", format(Sys.Date(), "%m-%Y.gz"))
geneID2pubmed_filename_txt = substr(geneID2pubmed_filename, 1, nchar(geneID2pubmed_filename)-3)
if(!file.exists(geneID2pubmed_filename)) {
  downloader::download(geneID2pubmed_url, geneID2pubmed_filename)
}
R.utils::gunzip(geneID2pubmed_filename, remove = F, overwrite = T)
# getting all pmids which have genes associated with them
geneID2pubmed = fread(geneID2pubmed_filename_txt, colClasses = c("character","character","character"))
geneID2pubmed_human = copy(geneID2pubmed[`#tax_id` == "9606",])
unlink(geneID2pubmed_filename_txt)
```

#### Map Entrez gene ids to UniprotKB ids

```{r map_Uniprot_ID_2_Entrez, message=F, warning=F}
uniprot = UniProt.ws(taxId=9606)
# grep("PUB",columns(unp), value =T)
human_genes_with_PMIDs = unique(geneID2pubmed_human[,GeneID])
entrezgene2uniprot = data.table()
for(i in seq(1, length(human_genes_with_PMIDs), 2000)){
        entrezgene2uniprot_temp = as.data.table(select(uniprot, 
               keys = human_genes_with_PMIDs[i:(i+2000)], 
               columns = c("ENTREZ_GENE", "UNIPROTKB"),
               keytype = "ENTREZ_GENE"))
        entrezgene2uniprot = rbind(entrezgene2uniprot, copy(entrezgene2uniprot_temp[complete.cases(entrezgene2uniprot_temp),]))
}
entrezgene2uniprot_not_in_imex = unique(entrezgene2uniprot[!UNIPROTKB %in% imex_human_proteins,])
```

 `r entrezgene2uniprot[, length(unique(UNIPROTKB))]` proteins (UNIPROTKB) have been mapped to genes.  
 `r entrezgene2uniprot[!UNIPROTKB %in% imex_human_proteins, length(unique(UNIPROTKB))]` proteins (UNIPROTKB) mapped to genes do not have interactions in IMEx.  
 `r entrezgene2uniprot[UNIPROTKB %in% imex_human_proteins, length(unique(UNIPROTKB))]` proteins (UNIPROTKB) mapped to genes do have interactions in IMEx.  

```{r save_pmids_with_proteins_witout_interactions}
NCBI_pmids = geneID2pubmed_human[GeneID %in% entrezgene2uniprot_not_in_imex$ENTREZ_GENE, unique(PubMed_ID)]
write(NCBI_pmids, "./results/NCBI_pmids_with_proteins_witout_interactions.txt")
```


#### the list of pmids associated with proteins without interactions can be accessed: https://github.com/vitkl/darkspaceproject/tree/master/dsp_comparison/results/NCBI_pmids_with_proteins_witout_interactions.txt


### protein to pmids mapping from UniProt

```{r protein_to_pmids_UniProt}
if(!file.exists("./source_files/protein_to_pmids_UniProt.txt.gz")){
        downloader::download("http://www.uniprot.org/uniprot/?compress=yes&query=organism:9606&format=tab&columns=id,organism-id,citation,citationmapping",
                             "./source_files/protein_to_pmids_UniProt.txt.gz")
        #R.utils::gzip("./source_files/protein_to_pmids_UniProt.txt")
}
R.utils::gunzip("./source_files/protein_to_pmids_UniProt.txt.gz", remove = F, overwrite = T)
protein_to_pmids_UniProt = fread("./source_files/protein_to_pmids_UniProt.txt", stringsAsFactors = F)
unlink("./source_files/protein_to_pmids_UniProt.txt")

colnames(protein_to_pmids_UniProt) = gsub(" ", "_", colnames(protein_to_pmids_UniProt))

UniProt_PubMed_ID = unique(protein_to_pmids_UniProt[!(Entry %in% imex_human_proteins) & PubMed_ID != "", unlist(strsplit(PubMed_ID, "; "))])
UniProt_Mapped_PubMed_ID = unique(protein_to_pmids_UniProt[!(Entry %in% imex_human_proteins) & Mapped_PubMed_ID != "", unlist(strsplit(Mapped_PubMed_ID, "; "))])
write(UniProt_PubMed_ID, "./results/UniProt_PubMed_ID_with_proteins_witout_interactions.txt")
write(UniProt_Mapped_PubMed_ID, "./results/UniProt_Mapped_PubMed_ID_with_proteins_witout_interactions.txt")
```

#### There are two columns in Uniprot and two resulting datasets:  

- PubMed_ID, which contains `r length(unique(protein_to_pmids_UniProt[PubMed_ID != "", unlist(strsplit(PubMed_ID, "; "))]))` publications associated with UniProtKB accession total, including `r length(UniProt_PubMed_ID)` publications associated with proteins with no interactions in IMEx.   
The dataset can be accessed: https://github.com/vitkl/darkspaceproject/tree/master/dsp_comparison/results/UniProt_PubMed_ID_with_proteins_witout_interactions.txt   
- Mapped_PubMed_ID, which contains `r length(unique(protein_to_pmids_UniProt[Mapped_PubMed_ID != "", unlist(strsplit(Mapped_PubMed_ID, "; "))]))` publications associated with UniProtKB accession total, including `r length(UniProt_Mapped_PubMed_ID)` publications associated with proteins with no interactions in IMEx.   
The dataset can be accessed: https://github.com/vitkl/darkspaceproject/tree/master/dsp_comparison/results/UniProt_Mapped_PubMed_ID_with_proteins_witout_interactions.txt  