IMEx dataset generator
========================================================

#### Libraries
```{r libraries,message=FALSE,warning=FALSE}
library(data.table)
```

#### Load IntAct data

```{r echo=FALSE, cache=TRUE}
intact_date <- date()
```

I download the latest version of the IntAct data. The file was downloaded on `r intact_date`. 

```{r}
if(!file.exists("./source_files/intact_mitab27.txt")){
  download.file("ftp://ftp.ebi.ac.uk/pub/databases/intact/current/psimitab/intact.txt", destfile = "./source_files/intact_mitab27.txt")
}
```

I need to clean up the intact file.

```{r eval=TRUE,message=FALSE, warning=FALSE}
system("perl ./scripts/MITAB27extractor_v13.pl ./source_files/intact_mitab27.txt ./processed_files/intact_pairs.txt")
```
```{r}
intact_full <- fread("./processed_files/intact_pairs.txt", header = T, sep = "\t", colClasses = "character")
```

#### Load data from legacy DIP

```{r eval=FALSE,echo=FALSE, cache=FALSE}
dip_date <- date()
```
I incorporate the data from legacy DIP as well, taking the latest available release there is. Data was downloaded on Mon Apr 11 16:52:07 2016 and needed to be downloaded manually, since the DIP download options only appear for logged-in users.  

Then I clean up the data as I did with the IntAct dataset. 

```{r eval=TRUE,message=FALSE, warning=FALSE}
system("perl ./scripts/MITAB25extractor_v12.pl ./source_files/dip20160731.txt ./processed_files/dip20160731_pairs.txt")
```
```{r}
dip_full <- fread("./processed_files/dip20160731_pairs.txt", header = T, sep = "\t", colClasses = "character")
```

#### Merge IntAct and DIP data to create an IMEx dataset

The original DIP data is in MITAB 2.5 format and IntAct is 2.7, so I select only 2.5-compatible files from the processed IntAct dataset. 

```{r eval=T}
intact_sel <- unique(subset(intact_full, select = c("pair_id", "id_a","id_b","pair_id_clean","id_a_clean","id_b_clean","taxid_a","taxid_b","pubid")))
```

Now I create a full IMEx dataset. 

```{r}
imex_full <- unique(rbind(intact_sel,dip_full))

write.table(imex_full, "./results/imex_full.txt", quote=F, sep ="\t", row.names = F, col.names = T)
```

The full dataset contains `r length(unique(imex_full$pair_id_clean))` interacting pairs (discarding isoforms and post-processed chains). Of these, `r length(unique(imex_full[taxid_a=="9606" & taxid_b=="9606"]$pair_id_clean))` are human pairs


I create a list of PMIDs that have been curated by IMEx-complying databases (IMPORTANT: not all these publications are curated to IMEx standards). 

```{r}
imex_pmids <- unique(imex_full[,.(pubid)])

write.table(imex_pmids, "./results/imex_pmids.txt", quote=F, sep ="\t", row.names = F, col.names = T)
```

There are `r nrow(imex_pmids)` publications in the full dataset of publications curated in IMEx-complying databases. Of these, `r nrow(unique(imex_full[taxid_a=="9606" & taxid_b=="9606",.(pubid)]))` are of human data only. 