---
title: Evaluation text mining predictions EPMC
date: 2020-01-17
author: Pablo Porras
---
```{r set-options, echo=FALSE}
options(width = 80)
```

IID interaction predictions: Technical report
========================================================

### Synopsis

I take the fraction of IID (http://dcv.uhnres.utoronto.ca/iid/) data that is either predicted using the FpClass algorithm (http://www.ncbi.nlm.nih.gov/pubmed/25402006) or derived from orthologous interactions as a predictor of potentially interacting pairs we have missed. 

#### Libraries
```{r libraries,message=FALSE,warning=FALSE}
library(data.table)
library(dplyr)
```

### Part 1: Download and process information from IID

I get the data from the IID web site.

```{r}
if(!file.exists("./source_files/human_annotated_PPIs.txt.gz")){
        download.file("http://iid.ophid.utoronto.ca/static/download/human_annotated_PPIs.txt.gz", destfile="./source_files/human_annotated_PPIs.txt.gz", method= "curl")
}

setwd("./source_files")
system("gunzip -d human_annotated_PPIs.txt.gz")
iidhumanall <- fread(
        "./human_annotated_PPIs.txt", 
        header = T, 
        sep = "\t",
        check.names = T)
system("rm human_annotated_PPIs.txt")
setwd("../")
```

I have a look at the data to check how much of it is 100% predicted. 

```{r comment=NA}
table(iidhumanall$evidence.type)
```

I will discard all the interactions that have been only detected experimentally. 

```{r comment=NA,warning=FALSE,message=FALSE}
iid_pred <- iidhumanall[evidence.type!="exp"]
table(iid_pred$evidence.type)
```

Now I select the fields I will need for the comparison. I keep the field in which IID codes the database of origin to assess if there are mapping issues in the comparison. 

```{r message=FALSE,warning=FALSE}
iid_pred_sel <- unique(select(iid_pred,uniprot1,uniprot2,dbs))
```

I generate pair identifiers for each interaction reported in the dataset.

```{r}
iid_pred_sel$pair_id <- apply(iid_pred_sel[,1:2], 1,function(i){
        paste(sort(i),collapse = "_")
})
iid_pred_sel$iid <- "yes"
```

I save this as a text file for further comparison with other datasets.

```{r}
iid_pred_pairs <- unique(select(iid_pred_sel,pair_id))
iid_pred_pairs$iid_pred <- 1
write.table(iid_pred_pairs,"./results//pairs_iid_pred.txt",col.names=T,row.names=F,quote=F,sep="\t")
```

The IID predicted dataset contains `r length(unique(iid_pred_pairs$pair_id))` unique pairs. 

### Part 2: Load in the IMEx dataset

I use a small pipeline to put together data from DIP and IntAct. The details can be found [here](https://github.com/pporrasebi/darkspaceproject/IMEx/IMEx_dsgen.md). 

```{r}
imex_full <- read.delim("../IMEx/results//imex_full.txt", header=T, sep="\t",colClasses="character")
```

### Part 3: IMEx-predicted IID comparison

I compare both sets and save the results for further use. 

```{r comment=NA}
imex_sel <- unique(select(imex_full,pair_id_clean,pair_id_clean,id_a_clean,id_b_clean,taxid_a,taxid_b,pubid))
imex_sel$imex <- "yes"
imex_pairs <- unique(select(imex_sel, pair_id=pair_id_clean,imex))

comp <- unique(merge(iid_pred_sel,imex_pairs,by="pair_id",all=T))

comp <- mutate(comp, db_pair =
                       ifelse(iid == "yes" & is.na(imex), "iid",
                       ifelse(is.na(iid) & imex == "yes", "imex",
                       ifelse(iid == "yes" & imex == "yes", "iid & imex",
                       "check"))))

comp$db_pair <- as.factor(comp$db_pair)

table(comp$db_pair,useNA="ifany")
```

I perform an additional check to see if there are interactions that IID obtained from IMEx-complying databases that were not properly mapped with my approach. 

```{r}

faulty_maps_imex <- unique(subset(comp, (grepl("intact",comp$dbs) | grepl("mint",comp$dbs) | grepl("dip",comp$dbs) | grepl("innatedb",comp$dbs)) & db_pair=="iid"))

```

There are `r nrow(faulty_maps_imex)` interacting pairs that should have been mapped to IMEx and were not, which represents `r round(nrow(faulty_maps_imex)*100/nrow(comp),2)`% of the total amount of pairs compared. I need to investigate further why this happens. For now, I will ignore it and generate a provisional file for further comparison with other approaches. 

```{r}
comp_simple <- unique(select(comp, pair_id,db_pair))

write.table(comp_simple,"./results/pairs_iid_vs_imex.txt",col.names=T,row.names=F,quote=F,sep="\t")
```