---
title: "DSP evaluation"
output: html_notebook
---

## Synopsis

Manual evaluation of Low-Hanging Fruit lists by curators in IMEx consortium. 

#### Libraries
```{r libraries, message=FALSE,warning=FALSE}
library(data.table)
```
### List upload

Lists are manually evaluated and compiled at https://docs.google.com/spreadsheets/d/1tL1HtVD3-BxHxKuXbIYhcFjmCptGVEOD5aFJCZw6CZk/edit#gid=0. Lists are manually downloaded from there. 

```{r upload}
# First LHF list
lhf_rd1 <- fread("./eval_tables/lhf_rd1.txt",header=T)

# EPMC text-mining lists
tm_it1_eval <- fread("./eval_tables/tm_it1.txt",header=T)
tm_it2_eval <- fread("./eval_tables/tm_it2.txt",header=T)
```

### Evaluation

#### LHF iteration 1

*Publications containing interactions*
```{r lhf_it1_eval_publ}
lhf_pub_pred <- unique(lhf_rd1[,.(pmid,`Contains interactions?`)])
lhf_pub_eval <- data.table(table(lhf_pub_pred$`Contains interactions?`))
```
Positives: `r paste(round(lhf_pub_eval[V1=="yes",.(N)]/sum(lhf_pub_eval[,.(N)])*100,2),"%",sep="")`  
Negatives: `r paste(round(lhf_pub_eval[V1=="no",.(N)]/sum(lhf_pub_eval[,.(N)])*100,2),"%",sep="")`


*Interacting protein pairs correctly identified*
```{r lhf_it1_eval_pairs}
lhf_pair_pred <- unique(lhf_rd1[,.(pair_id,`Are pairs correct?`)])
lhf_pair_eval <- data.table(table(lhf_pair_pred$`Are pairs correct?`))
```
Positives: `r paste(round(lhf_pair_eval[V1=="yes",.(N)]/sum(lhf_pair_eval[,.(N)])*100,2),"%",sep="")`  
Negatives: `r paste(round(lhf_pair_eval[V1=="no",.(N)]/sum(lhf_pair_eval[,.(N)])*100,2),"%",sep="")`

*Correct new pairs*
```{r new pairs}
lhf_newpairs <- unique(lhf_rd1[noncur_prot==1,.(pair_id,`Are pairs correct?`)])
lhf_newpairs_eval <- data.table(table(lhf_newpairs$`Are pairs correct?`))
newpairsfound <- lhf_newpairs_eval[V1=="yes",.(N)]
newpairsfound
```