---
title: "DSP lists to coordinate curation"
author: "Pablo Porras"
date: "05/12/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,eval=FALSE)
```

## Summary

This is an update for the set of scripts to generate a final curation list that can be used by IMEx curators in order to prioritize paper selection. 

#### Libraries
```{r libraries,message=FALSE,warning=FALSE}
library(data.table)
library(googlesheets)
library(splitstackshape)
library(RCurl)
library(jsonlite)
```

### Upload lists
#### Tested publications known to contain interaction data

As downloaded on `r date()`.
```{r tested,message=FALSE,warning=FALSE}
# gs_auth()
# key <- extract_key_from_url("https://docs.google.com/spreadsheets/d/1tL1HtVD3-BxHxKuXbIYhcFjmCptGVEOD5aFJCZw6CZk/edit#gid=989290982")
# dsp_manual <- gs_key(key)
# pmids_checked <- data.table(gs_read(ss=dsp_manual, ws = "dsp_priority",header=T))
pmids_checked <- fread(
        "./source_data/Evaluating DSP - dsp_priority_rev.tsv",
        check.names = T,
        sep = "\t",
        colClasses = "character",
        stringsAsFactors = F
)

pmids_checked$pmid <- as.character(pmids_checked$pmid)
pmids_checked <- pmids_checked[,sort:=.I]
```

#### Ranked list of DSP publications
As provided by Miguel Vázquez on 16/10/2018. 
```{r rank,message=F,warning=F}
dsp_rank <- fread(input = "zcat < ../../DarkSpace/manual_evaluation/round5/pmid_ranks.gz",
                  header=T,
                  sep="\t",
                  check.names = T,
                  stringsAsFactors = F)

dsp_rank_sel <- unique(dsp_rank[,.(pmid = as.character(X.PMID),
                                   relevance.up = round(Relevance,2),
                                   DS.interest.up = Dark.Space.interest,
                                   DS.part_interest.up = Dark.Space..partial..interest)])
```


#### Extracting information from the EPMC API to annotate publications

Steps: 

 - Map PMIDs to PMC ids
 - Obtain MESH terms annotations from EPMC API.
        - Can be extracted using this example: `https://www.ebi.ac.uk/europepmc/webservices/rest/search?query=ext_id:18074021%20AND%20%28PUB_TYPE:Research-article%20NOT%20PUB_TYPE:Review%20NOT%20PUB_TYPE:review-article%29&resultType=core&format=json`
        - Fields to extract: meshHeading. 
 - Obtain mutation information from PubTator
        - Up to 8 ids per request
        - Request example: `https://www.ebi.ac.uk/europepmc/annotations_api/annotationsByArticleIds?articleIds=MED%3A30497962%2CMED%3A24194929&type=Gene%20Mutations&provider=PubTator_NCBI&format=JSON`.  


##### MESH terms        
```{r api_query_kw_mesh}
# set.seed(88)
# pmid_query <- pmids_checked[sample.int(nrow(pmids_checked),100)]$pmid

pmid_query <- pmids_checked$pmid

pmid_mesh_kw <- data.frame(
        pmid = character(),
        mesh = character(),
        reqanswer = character(),
        review = character(),
        stringsAsFactors = F)

for (i in 1:length(pmid_query)){

        pmid_mesh_kw[i,]$pmid = pmid_query[i]
        
        url = paste("https://www.ebi.ac.uk/europepmc/webservices/rest/search?query=ext_id:",pmid_query[i],"%20AND%20%28PUB_TYPE:Research-article%20NOT%20PUB_TYPE:Review%20NOT%20PUB_TYPE:review-article%29&resultType=core&format=json",sep="")
        urldetails <- getURL(url,header=T)
        header <- gsub("\r\nServer:.*","",urldetails)
        
        pmid_mesh_kw[i,]$reqanswer = header
        
        if(grepl("200",header)){
                jsonraw <- gsub(".*\\{\"version\"","\\{\"version\"",urldetails)
                json <- fromJSON(jsonraw)
                
                temp_mesh <- data.table((json$resultList$result$meshHeadingList$meshHeading[[1]]))
                
                pmid_mesh_kw[i,]$mesh <- paste(unique(temp_mesh$descriptorName),collapse="|")
                pmid_mesh_kw[i,]$review <- ifelse(
                        !is.null(json$resultList$result$id),
                        "no",
                        "yes"
                )
                
        }
        
        else{
                pmid_mesh_kw[i,]$mesh <- ""
        }
}
```

##### PubTator
```{r api_mut}
pmid_pubtator_mut <- data.frame(
        pmid = character(),
        mutannot = character(),
        reqanswer = character(),
        stringsAsFactors = F)

for (i in 1:length(pmid_query)){
        
        pmid_pubtator_mut[i,]$pmid = pmid_query[i]
        
        url = paste("https://www.ebi.ac.uk/europepmc/annotations_api/annotationsByArticleIds?articleIds=MED%3A",pmid_query[i],"&type=Gene%20Mutations&provider=PubTator_NCBI&format=JSON",sep="")
        urldetails <- getURL(url,header=T)
        header <- gsub("\r\nServer:.*","",urldetails)
        
        pmid_pubtator_mut[i,]$reqanswer = header
        
        if(grepl("200",header)){
                jsonraw <- gsub("]$","",gsub(".*\\{\"source\"","\\{\"source\"",urldetails))
                json <- fromJSON(jsonraw)
                
                pmid_pubtator_mut[i,]$mutannot <- ifelse(
                        is.null(json$annotations[1][[1]]),
                        "no",
                        "yes"
                )
        }
        
        else{
                pmid_pubtator_mut[i,]$mutannot <- ""
        }
}

```
 

### Putting all lists together
```{r merge}
pre_dsp_rev <- unique(merge(
        pmids_checked,
        pmid_mesh_kw,
        by = "pmid",
        all.x = T,
        all.y = T
))

pre_dsp_rev2 <- unique(merge(
        pre_dsp_rev,
        pmid_pubtator_mut,
        by = "pmid",
        all = T
))

pre_dsp_rev3 <- unique(merge(
        pre_dsp_rev2,
        dsp_rank_sel,
        by = "pmid",
        all.x = T,
        all.y = F
))

dsp_rev <- pre_dsp_rev3[order(sort),.(
        pmid,
        curator,
        curatable,
        curator_comment,
        prot_pairs,
        mesh,
        alzheimer = ifelse(                             ### This was added to identify Alzheimer's 
                grepl("Alzheimer|alzheimer",mesh),      ### disease-related publications for a  
                "yes",                                  ### project in UniProt. 
                "no"
        ),
        review,
        mutannot,
        relevance.up,
        DS.interest.up,
        DS.part_interest.up,
        relevance,
        DS.interest,
        DS.part_interest,
        ccline_uniprot,
        noncur_prot,
        contains_ints_man_check,
        contains_ints_man_check_comment,
        imex_publ,
        imex_pair,
        GO_IPI_publ,
        GO_IPI_pair,
        BioGRID_publ,
        BioGRID_pair,
        tm_epmc_publ,
        tm_epmc_pair,
        EVEX_publ,
        EVEX_pair,
        STRING_textmining_pair,
        reactome_publ,
        reactome_pair,
        OmniPath_interactions_publ,
        OmniPath_interactions_pair,
        OmniPath_ptm_publ,
        OmniPath_ptm_pair,
        STRING_pi_pair,
        iid_pred
)]

fwrite(
        dsp_rev,
        "./results/dsp_priority_rev.txt",
        col.names = T,
        row.names = F,
        sep="\t",
        quote=F)
```

This table is uploaded into our DSP google spreadsheet at https://docs.google.com/spreadsheets/d/1tL1HtVD3-BxHxKuXbIYhcFjmCptGVEOD5aFJCZw6CZk/edit#gid=826915149. We only upload the first 10,000 rows from the priority table, to avoid size limitation issues in google docs. 
