---
title: "DSP lists to coordinate curation: update adding TM method and STRING predictions"
author: "Pablo Porras"
date: "03/04/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,eval=FALSE)
```

## Summary

This is an update for the set of scripts to generate a final curation list that can be used by IMEx curators in order to prioritize paper selection. This update just adds two columns bearing the text-mined experimental methods as extracted by EMPC pipeline and a column stating if the pair has been predicted by STRING using phylogeny. 

#### Libraries
```{r libraries,message=FALSE,warning=FALSE}
library(data.table)
library(googlesheets)
```

### Upload lists
#### DSP subset actively curated in Google docs
As downloaded on `r date()`.
```{r tested,message=FALSE,warning=FALSE}
# This code does not work due to some issue with the permissions given by the app used. 
# dsp_manual <- gs_title("Evaluating DSP")
# pmids_checked <- data.table(gs_read(ss=dsp_manual, ws = "All_PMIDs_checked",header=T))
# pmids_checked$pmid <- as.character(pmids_checked$pmid)
# I need to download the data manually and then upload it here
pmids_checked <- unique(fread(
        "./source_data/20200305.Evaluating DSP - All_PMIDs_checked.tsv",
        sep = "\t",
        header = T,
        check.names = T,
        stringsAsFactors = F,
        colClasses = "character"
))
```

#### Full DSP pair-publication list
```{r dsp_comp}
setwd("../dsp_comparison/results/")
system("tar -zxvf pubpaircomp_table_final.txt.tar.gz")
tm_pairs_pmids <- fread(
        "pubpaircomp_table_final.txt",
        header=T,
        sep="\t",
        colClasses="character",
        check.names = T,
        stringsAsFactors = F)
system("rm pubpaircomp_table_final.txt")
setwd("../../dsp_lists/")
```

#### Select only relevant fields of DSP pair-publication list
```{r dsp_wrangling}
tm_pmids_method_pred <- unique(tm_pairs_pmids[,.(
        tm_method_term = paste(unique(tm_method_term),collapse = "|"),
        STRING_phylo = paste(unique(STRING_phylo),collapse = "|")
), by = "pmid"])

tm_pmids_method_pred <- tm_pmids_method_pred[,.(
        pmid,
        tm_method_term = gsub("0\\||\\|0","",tm_method_term),
        STRING_phylo = ifelse(
                grepl("1",STRING_phylo),
                "yes",
                "no"
        )
)]
```

#### Update DSP curated list
```{r dsp_curapted_update}
dsp_curation_updated <- unique(merge(
        dsp_curation,
        tm_pmids_method_pred,
        all.x = T,
        all.y = F
))

dsp_curation_updated <- dsp_curation_updated[,.(
        pmid,curator,curatable,curator_comment,contains_ints_man_check,contains_ints_man_check_comment,prot_pairs,mesh,alzheimer,review,mutannot,relevance.up,DS.score.pred,DS.interest.up,DS.part_interest.up,relevance,DS.interest,DS.part_interest,ccline_uniprot,noncur_prot,imex_publ,imex_pair,GO_IPI_publ,GO_IPI_pair,BioGRID_publ,BioGRID_pair,tm_epmc_publ,tm_epmc_pair,tm_method_term,EVEX_publ,EVEX_pair,STRING_textmining_pair,reactome_publ,reactome_pair,OmniPath_interactions_publ,OmniPath_interactions_pair,OmniPath_ptm_publ,OmniPath_ptm_pair,STRING_pi_pair,iid_pred,STRING_phylo,sort
)]

fwrite(
        dsp_curation_updated,
        "./results/dsp_curation_updated.txt",
        col.names = T,
        row.names = F,
        quote = F,
        sep = "\t"
)
```

This table is uploaded into our DSP google spreadsheet at https://docs.google.com/spreadsheets/d/1tL1HtVD3-BxHxKuXbIYhcFjmCptGVEOD5aFJCZw6CZk/edit#gid=826915149. We only upload the first 10,000 rows from the priority table, to avoid size limitation issues in google docs. 
