---
title: Reactome dataset generation
date: 2020-01-17
author: Pablo Porras
---
```{r set-options, echo=FALSE}
options(width = 80)
```
Generation of the Reactome dataset: Technical report
========================================================

### Synopsis

The goal of this analysis is to use Reactome to try to find potentially non-curated protein interactions for which there is experimental evidence. I define a set of associated pairs as found in Reactome. Those proteins that are found to be present in the same reaction are combined according to their role.  

#### Libraries
```{r libraries,message=FALSE,warning=FALSE}
library(data.table)
library(UniProt.ws)
library(splitstackshape)
```

### Part 1: Loading Reactome data

Files downloaded from Reactome. As a pre-load operation, I clean up the UniProtKB accessions given by Reactome to avoid using isoform identifiers. 

```{r message=FALSE,warning=FALSE}
if(!file.exists("./source_files/reactome.homo_sapiens.interactions.psi-mitab.txt")){
  download.file("https://reactome.org/download/current/interactors/reactome.homo_sapiens.interactions.psi-mitab.txt", destfile = "./source_files/reactome.homo_sapiens.interactions.psi-mitab.txt")
}

homo_sapiens_interactions = fread(
        "./source_files/reactome.homo_sapiens.interactions.psi-mitab.txt",
        colClasses = "character",
        sep = "\t", 
        quote = "")

# view which types of interactions are present
homo_sapiens_interactions[,table(`Interaction type(s)`)]
homo_sapiens_interactions[,table(`Host organism(s)`)]
# identify which interacting partners 
homo_sapiens_interactions[,table(gsub(":.+$","",unique(c(`#ID(s) interactor A`,`ID(s) interactor B`))))]

# clean interactor ids from prefix and isoform id and filter out non-uniprot pairs
homo_sapiens_interactions = homo_sapiens_interactions[grep("uniprotkb",`#ID(s) interactor A`),]
homo_sapiens_interactions = homo_sapiens_interactions[grep("uniprotkb",`ID(s) interactor B`),]
homo_sapiens_interactions[,`#ID(s) interactor A` := gsub("\\-.+$","",gsub("^uniprotkb:","",`#ID(s) interactor A`))]
homo_sapiens_interactions[,`ID(s) interactor B` := gsub("\\-.+$","",gsub("^uniprotkb:","",`ID(s) interactor B`))]
```

### Part 2: Select only human proteins

I need to query uniprot via Uniprot.ws package which provides easy access to Uniprot web sevices, search is limited by species, so I woundn't find matches for non-human species represented in the dataset and can use this information to filter out interaction containing non-human proteins. 

```{r}
proteins = homo_sapiens_interactions[,unique(c(`#ID(s) interactor A`,`ID(s) interactor B`))]

write(proteins, "./processed_files/protein_updated_reactome.txt")

 uniprot = UniProt.ws(taxId=9606)
# getting names of the keys and columns from Uniprot
# keytypes(uniprot)
# columns(uniprot)
# mapping IDs
 proteins2species <- as.data.table(select(uniprot, 
               keys = proteins, 
               columns = c("ORGANISM-ID", "UNIPROTKB"),
               keytype = "UNIPROTKB"))
 # Filter non-mapped proteins (e.g. non-human)
 proteins2species = copy(proteins2species[complete.cases(proteins2species),])
 
# filter non-human only interactions:
homo_sapiens_interactions = homo_sapiens_interactions[`#ID(s) interactor A` %in% proteins2species$UNIPROTKB,][`ID(s) interactor B` %in% proteins2species$UNIPROTKB,]
```

### Part 3: Generate Reactome association pairs

I finally add the PMID information to each pair of associations and I check if there are reactions for which no source PMID is available. Those are identified because they have the Reactome publication as reference (pubmed:24243840).

```{r }
homo_sapiens_interactions[, pmids := gsub("pubmed:","",`Publication Identifier(s)`)]

homo_sapiens_interactions2 <- homo_sapiens_interactions[pmids!="24243840"]

homo_sapiens_interactions2[, pair_id := apply(data.table(`#ID(s) interactor A`,`ID(s) interactor B`,stringsAsFactors = F), 1,
                                               function(a) { z = sort(a)
                                               paste0(z[1],"_",z[2]) })]

homo_sapiens_interactions3 <- cSplit(
        homo_sapiens_interactions2,
        splitCols = "pmids",
        sep = "|",
        drop = F,
        direction = "long"
)

homo_sapiens_interactions_final <- unique(homo_sapiens_interactions3[,.(
        pair_id,
        pmid = pmids,
        reactome = 1
)])
```
`r length(unique(homo_sapiens_interactions_final$pair_id))` interactions have associated pmids. There is a total of `r length(unique(homo_sapiens_interactions_final$pmid))` publications in the data set.

### Part 4: Final formatting steps for generation of the Reactome dataset

I generate a unique pair identifier per line and save all Reactome pairs including those associated with the PMIDs as a text file for further comparison with other datasets.

```{r message=FALSE,warning=FALSE}
fwrite(homo_sapiens_interactions_final[,.(pair_id, pmid, reactome)], "./results/pairs_pmid_reactome.txt", sep = "\t")
system("gzip ./results/pairs_pmid_reactome.txt --force")
```

After collapsing redundant pairs, I end up with a list of `r nrow(homo_sapiens_interactions_final)` unique protein pair-publication associations extracted from Reactome in human. 
