---
title: Reactome dataset generation
date: 2016-04-13
author: Pablo Porras
---
```{r set-options, echo=FALSE}
options(width = 80)
```
Generation of the Reactome dataset: Technical report
========================================================

### Synopsis

The goal of this analysis is to use Reactome to try to find potentially non-curated protein interactions for which there is experimental evidence. I define a set of associated pairs as found in Reactome. Those proteins that are found to be present in the same reaction are combined according to their role.  

### Part 1: Loading Reactome data

The files used for this analysis were supplied by Florent Yvon at Reactome around end of April, 2017. I cannot use the interactions as exported by Reactome since they do not give a PMID for each inferred interacting pair and are then not usable. As a pre-load operation, I clean up the UniProtKB accessions given by Reactome to avoid using isoform identifiers. 

```{r message=FALSE,warning=FALSE}
library(data.table)
library(UniProt.ws)
homo_sapiens_interactions = fread("./source_files/homo_sapiens_interactions.txt", colClasses = "character",sep = "\t", quote = "")

# view which types of interactions are present
homo_sapiens_interactions[,table(`Interaction type(s)`)]
homo_sapiens_interactions[,table(`Host organism(s)`)]
# identify which interacting partners 
homo_sapiens_interactions[,table(gsub(":.+$","",unique(c(`#ID(s) interactor A`,`ID(s) interactor B`))))]

# clean interactor ids from prefix and isoform id and filter out non-uniprot pairs
homo_sapiens_interactions = homo_sapiens_interactions[grep("uniprotkb",`#ID(s) interactor A`),]
homo_sapiens_interactions = homo_sapiens_interactions[grep("uniprotkb",`ID(s) interactor B`),]
homo_sapiens_interactions[,`#ID(s) interactor A` := gsub("\\-.+$","",gsub("^uniprotkb:","",`#ID(s) interactor A`))]
homo_sapiens_interactions[,`ID(s) interactor B` := gsub("\\-.+$","",gsub("^uniprotkb:","",`ID(s) interactor B`))]
```

### Part 2: Select only human proteins

I need to query uniprot via Uniprot.ws package which provides easy access to Uniprot web sevices, search is limited by species, so I woundn't find matches for non-human species represented in the dataset and can use this information to filter out interaction containing non-human proteins. 

```{r}
proteins = homo_sapiens_interactions[,unique(c(`#ID(s) interactor A`,`ID(s) interactor B`))]

write(proteins, "./processed_files/protein_updated_reactome.txt")

 uniprot = UniProt.ws(taxId=9606)
# getting names of the keys and columns from Uniprot
# keytypes(uniprot)
# columns(uniprot)
# mapping IDs
 proteins2species <- as.data.table(select(uniprot, 
               keys = proteins, 
               columns = c("ORGANISM-ID", "UNIPROTKB"),
               keytype = "UNIPROTKB"))
 # Filter non-mapped proteins (e.g. non-human)
 proteins2species = copy(proteins2species[complete.cases(proteins2species),])
 
# filter non-human only interactions:
homo_sapiens_interactions = homo_sapiens_interactions[`#ID(s) interactor A` %in% proteins2species$UNIPROTKB,][`ID(s) interactor B` %in% proteins2species$UNIPROTKB,]
```

### Part 3: Generate Reactome association pairs

I finally add the PMID information to each pair of associations and I check if there are reactions for which no PMID is available. 

```{r }
homo_sapiens_interactions[, pmid := ""]
homo_sapiens_interactions[grep("pubmed",`Interaction Xref(s)`), pmid := grep("pubmed",`Interaction Xref(s)`, value = T)]

homo_sapiens_interactions[, pair_id := apply(data.table(`#ID(s) interactor A`,`ID(s) interactor B`,stringsAsFactors = F), 1,
                                               function(a) { z = sort(a)
                                               paste0(z[1],"_",z[2]) })]

homo_sapiens_interactions2 = copy(homo_sapiens_interactions[pmid != "",])

homo_sapiens_interactions2 = copy(homo_sapiens_interactions2[, unlist(gsubfn::strapplyc(pmid,"pubmed:[[:digit:]]+",simplify = T)), by = pair_id])
homo_sapiens_interactions2[,reactome := 1]
colnames(homo_sapiens_interactions2)[2] = "pmid"
homo_sapiens_interactions2[, pmid := gsub("pubmed:","",pmid)]

homo_sapiens_interactions3 = unique(rbind(homo_sapiens_interactions2, homo_sapiens_interactions[pmid == "", .(pair_id, pmid, reactome =1)]))
```
`r homo_sapiens_interactions[, sum(pmid != "")]` interactions have associated pmids, `r homo_sapiens_interactions[, sum(pmid == "")]` do not have associated pmids.

### Part 4: Final formatting steps for generation of the Reactome dataset

I generate a unique pair identifier per line and save all Reactome pairs including those associated with the PMIDs as a text file for further comparison with other datasets.

```{r message=FALSE,warning=FALSE}
fwrite(homo_sapiens_interactions3[,.(pair_id, pmid, reactome)], "./results/pairs_pmid_reactome.txt", sep = "\t")
system("gzip ./results/pairs_pmid_reactome.txt --force")
```

After collapsing redundant pairs, I end up with a list of `r nrow(homo_sapiens_interactions3)` unique protein pair-publication associations extracted from Reactome in human. 