---
title: "DSP evaluation pairs MIscore"
author: "Pablo Porras"
date: "14/02/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Evaluation of checked and curated publications of the DSP

I manually download all DSP evaluated PMIDs and also those already curated and avaliable on the IntAct website. 

##### Libraries
```{r libraries, message=FALSE,warning=FALSE}
library(data.table)
library(splitstackshape)
library(stringr)
library(ggplot2)
library(scales)
library(dplyr)
library(ggsignif)
```

#### Data upload
```{r upload}
# These come from the Google spreadsheet, manually downloaded. 
eval_dsp <- fread("./source_data/20190214_dsp_priority_rev_pred.tsv",
                  header = T,
                  colClasses = "character",
                  stringsAsFactors = F,
                  check.names = T,
                  sep = "\t")

if (!file.exists("./source_data/intact.zip")){
        download.file(url="ftp://ftp.ebi.ac.uk/pub/databases/intact/current/psimitab/intact.zip",destfile = "./source_data/intact.zip",method = "curl")
}

if (!file.exists("./source_data/intact-micluster.zip")){
        download.file(url="ftp://ftp.ebi.ac.uk/pub/databases/intact/current/psimitab/intact-micluster.zip",destfile = "./source_data/intact-micluster.zip",method = "curl")
}


setwd("./source_data")
system("unzip -j intact.zip")
intact <- unique(fread(
  "intact.txt",
  header=T,
  sep="\t",
  colClasses="character",
  check.names = T,
  stringsAsFactors = F))
system("rm intact.txt")
system("rm intact_negative.txt")

system("unzip -j intact-micluster.zip")
intact_micluster <- unique(fread(
  "intact-micluster.txt",
  header=T,
  sep="\t",
  colClasses="character",
  check.names = T,
  stringsAsFactors = F))
system("rm intact-micluster.txt")
system("rm intact-micluster_negative.txt")
setwd("../")
```

### Extract comparison sets
```{r comp_sets}
eval_dsp_pmids <- unique(eval_dsp[grepl("yes|Yes",curatable),.(pmid)])

# fwrite(
#   eval_dsp_pmids,
#   "./source_data/pmids.txt",
#   col.names = F,
#   row.names = F,
#   sep = "\t",
#   quote = F
# )

eval_dsp_pairs <- unique(eval_dsp[grepl("yes|Yes",curatable),.(
  prot_pairs,
  pmid,
  DS.interest.up,
  DS.part_interest.up,
  DS.score.pred)])

eval_dsp_pairs_long <- unique(cSplit(
  eval_dsp_pairs,
  splitCols = "prot_pairs",
  sep = "|",
  direction = "long",
  drop = F
))

eval_dsp_pairs_formatted <- eval_dsp_pairs_long[,.(
  pmids = paste(unique(pmid),collapse = "|"),
  p1 = gsub("_.*", "",prot_pairs),
  p2 = gsub(".*_", "",prot_pairs),
  DS.interest.up,
  DS.part_interest.up,
  DS.score.pred
), by = "prot_pairs"]

intact_human <- intact[grepl('taxid:9606',Taxid.interactor.A) &
                         grepl('taxid:9606',Taxid.interactor.B) &
                         grepl('uniprotkb:',X.ID.s..interactor.A) &
                         grepl('uniprotkb:',ID.s..interactor.B),]
intact_human <- intact_human[,pmid:=gsub("\\|.*","",gsub(".*pubmed:","",Publication.Identifier.s.))]
intact_human <- intact_human[,p1:=gsub("-.*","",gsub("uniprotkb:","",X.ID.s..interactor.A))]
intact_human <- intact_human[,p2:=gsub("-.*","",gsub("uniprotkb:","",ID.s..interactor.B))]
intact_human <- intact_human[,pair_id := apply(data.table(p1,p2,stringsAsFactors = F),
                  1,
                  function(a) { z = sort(a)
                  paste0(z[1],"_",z[2]) })]

intact_human <- data.table(mutate(intact_human,
                                  X.ID.s..interactor.A = paste("uniprotkb:",p1,sep=""),
                                  ID.s..interactor.B = paste("uniprotkb:",p2,sep="")))

publs_curated_dsp <- intact_human[pmid %in% eval_dsp_pmids$pmid]

intact_dsp <- intact_human[(pmid %in% eval_dsp_pmids$pmid)]
dsp_pairs_in_intact <- unique(intact_dsp[,.(
  p1,
  p2,
  pair_id = apply(data.table(p1,p2,stringsAsFactors = F),
                  1,
                  function(a) { z = sort(a)
                  paste0(z[1],"_",z[2]) })
)])

intact_pairs_dsp <- intact_human[(p1 %in% dsp_pairs_in_intact$p1 & p2 %in% dsp_pairs_in_intact$p2) | (p1 %in% dsp_pairs_in_intact$p2 & p2 %in% dsp_pairs_in_intact$p1)]

intact_pairs_dsp_no_dsppmids <- intact_pairs_dsp[!(pmid %in% eval_dsp_pmids$pmid)]

intact_pairs_dsp_mitab <- intact_pairs_dsp[ ,`:=`(pmid = NULL, p1 = NULL, p2 = NULL, pair_id = NULL)]
fwrite(
  intact_pairs_dsp_mitab,
  "./processed_data/intact_pairs_dsp_mitab.txt",
  col.names = T,
  row.names = F,
  quote = F,
  sep = "\t"
)

intact_pairs_dsp_no_dsppmids_mitab <- intact_pairs_dsp_no_dsppmids[ ,`:=`(pmid = NULL, p1 = NULL, p2 = NULL, pair_id = NULL)]
fwrite(
  intact_pairs_dsp_no_dsppmids_mitab,
  "./processed_data/intact_pairs_dsp_no_dsppmids_mitab.txt",
  col.names = T,
  row.names = F,
  quote = F,
  sep = "\t"
)
```

#### Obtain non-DSP publication sets for comparison
I need to have a set publications unrelated with the DSP, for which I run the same comparison as I do for the DSP publications. I select these among low-through put publications, checking what is the maximum amount of expanded interactions that DSP publications have. I decide to produce two sets, one that covered the time frame during which the DSP has been active and another predating it. The reason for this is that Margaret was specifically targeting proteins with no UniProt CC lines during the time of the DSP, so it could appear as if the 'control' set is actually better than the DSP at finding novel interacting pairs. 

##### CC lines subset
```{r random_cclines}
# Count expanded interactions
publ_count_exp_int <- unique(intact_human[,.(
  nr_exp_ints = .N,
  Creation.date = as.Date(Creation.date, "%Y/%m/%d"),
  Source.database.s.
),
by = "pmid"])

# Map to DSP publications
eval_dsp_pmids_count <- unique(merge(
  eval_dsp_pmids,
  publ_count_exp_int,
  by = "pmid",
  all.x = T,
  all.y = F
))

# Max number of expanded binaries per DSP publication
max(eval_dsp_pmids_count[!is.na(nr_exp_ints)]$nr_exp_ints)


# Select low-throughput publications only from intact_human
lt_publs <- publ_count_exp_int[nr_exp_ints < 100]

# Select the subset of publications, roughly within the same period as DSP, which do not include DSP pairs and excluding DIP records, since they come from mass import and the creation date is not indicative of when they were curated. 
lt_publs_no_dsp_year <- lt_publs[Creation.date >= min(eval_dsp_pmids_count[!is.na(nr_exp_ints)]$Creation.date) &
                                   Creation.date <= max(eval_dsp_pmids_count[!is.na(nr_exp_ints)]$Creation.date) &
                                   !(pmid %in% eval_dsp_pmids$pmid) &
                                   Source.database.s. != 'psi-mi:"MI:0465"(DIP)']

# Select protein pairs from these publications
no_dsp_pairs <- unique(intact_human[pmid %in% lt_publs_no_dsp_year$pmid &
                                      !is.na(p1) &
                                      !is.na(p2),.(
  p1,
  p2,
  pair_id = apply(data.table(p1,p2,stringsAsFactors = F),
                  1,
                  function(a) { z = sort(a)
                  paste0(z[1],"_",z[2]) })
)])

# Produce adequate MITAB files for non DSP publications selected. REMEMBER TO PUT A HASH ON THE COLUMN TITLES!!!
intact_pairs_nodsp <- intact_human[((p1 %in% no_dsp_pairs$p1 & p2 %in% no_dsp_pairs$p2) | (p1 %in% no_dsp_pairs$p2 & p2 %in% no_dsp_pairs$p1)) & Source.database.s. != 'psi-mi:"MI:0465"(DIP)']
intact_pairs_nodsp_no_dsppmids <- intact_pairs_nodsp[!(pmid %in% lt_publs_no_dsp_year$pmid)]

intact_pairs_nodsp_mitab <- intact_pairs_nodsp[ ,`:=`(pmid = NULL, p1 = NULL, p2 = NULL, pair_id = NULL)]
fwrite(
  intact_pairs_nodsp_mitab,
  "./processed_data/intact_pairs_nodsp_mitab.txt",
  col.names = T,
  row.names = F,
  quote = F,
  sep = "\t"
)

intact_pairs_nodsp_no_dsppmids_mitab <- intact_pairs_nodsp_no_dsppmids[ ,`:=`(pmid = NULL, p1 = NULL, p2 = NULL, pair_id = NULL)]
fwrite(
  intact_pairs_nodsp_no_dsppmids_mitab,
  "./processed_data/intact_pairs_nodsp_no_dsppmids_mitab.txt",
  col.names = T,
  row.names = F,
  quote = F,
  sep = "\t"
)
```

##### Old random subset
```{r random_old}
# Select the subset of publications, well before DSP, which do not include DSP pairs and excluding MINT records, since they come from mass import and the creation date is not indicative of when they were curated. 
lt_publs_no_dsp_old_year <- lt_publs[Creation.date >= "2014-06-01" &
                                   Creation.date <= "2016-01-21" &
                                   !(pmid %in% eval_dsp_pmids$pmid) &
                                   grepl("MINT",Source.database.s.)]

# Select protein pairs from these publications
no_dsp_old_pairs <- unique(intact_human[pmid %in% lt_publs_no_dsp_old_year$pmid &
                                      !is.na(p1) &
                                      !is.na(p2),.(
  p1,
  p2,
  pair_id = apply(data.table(p1,p2,stringsAsFactors = F),
                  1,
                  function(a) { z = sort(a)
                  paste0(z[1],"_",z[2]) })
)])

# Produce adequate MITAB files for non DSP publications selected. REMEMBER TO PUT A HASH ON THE COLUMN TITLES!!!
intact_pairs_nodsp_old <- intact_human[((p1 %in% no_dsp_old_pairs$p1 & p2 %in% no_dsp_old_pairs$p2) | (p1 %in% no_dsp_old_pairs$p2 & p2 %in% no_dsp_old_pairs$p1)) &  grepl("MINT",Source.database.s.)]
intact_pairs_nodsp_old_no_dsppmids <- intact_pairs_nodsp_old[!(pmid %in% lt_publs_no_dsp_old_year$pmid)]

intact_pairs_nodsp_old_mitab <- intact_pairs_nodsp_old[ ,`:=`(pmid = NULL, p1 = NULL, p2 = NULL, pair_id = NULL)]
fwrite(
  intact_pairs_nodsp_old_mitab,
  "./processed_data/intact_pairs_nodsp_old_mitab.txt",
  col.names = T,
  row.names = F,
  quote = F,
  sep = "\t"
)

intact_pairs_nodsp_old_no_dsppmids_mitab <- intact_pairs_nodsp_old_no_dsppmids[ ,`:=`(pmid = NULL, p1 = NULL, p2 = NULL, pair_id = NULL)]
fwrite(
  intact_pairs_nodsp_old_no_dsppmids_mitab,
  "./processed_data/intact_pairs_nodsp_old_no_dsppmids_mitab.txt",
  col.names = T,
  row.names = F,
  quote = F,
  sep = "\t"
)
```

### Uploading MIscore before/after
I get the MIscore values applying the algorithm on the local files produced in the previous step. REMEMBER TO PUT A HASH ON THE COLUMN TITLES!!!
```{r miscore_upload}
intact_pairs_dsp_mitab_miscore <- fread(
  "./processed_data/intact_pairs_dsp_clustered.txt",
  header = F,
  check.names = T,
  stringsAsFactors = F,
  colClasses = "character"
)
colnames(intact_pairs_dsp_mitab_miscore) <- colnames(intact_pairs_dsp_no_dsppmids_mitab)[1:15]

intact_pairs_dsp_no_dsppmids_mitab_miscore <- fread(
  "./processed_data/intact_pairs_dsp_no_dsppmids_clustered.txt",
  header = F,
  check.names = T,
  stringsAsFactors = F,
  colClasses = "character"
)
colnames(intact_pairs_dsp_no_dsppmids_mitab_miscore) <- colnames(intact_pairs_dsp_no_dsppmids_mitab)[1:15]

intact_pairs_nodsp_mitab_miscore <- fread(
  "./processed_data/intact_pairs_nodsp_clustered.txt",
  header = F,
  check.names = T,
  stringsAsFactors = F,
  colClasses = "character"
)
colnames(intact_pairs_nodsp_mitab_miscore) <- colnames(intact_pairs_dsp_no_dsppmids_mitab)[1:15]

intact_pairs_nodsp_no_dsppmids_mitab_miscore <- fread(
  "./processed_data/intact_pairs_nodsp_no_dsppmids_clustered.txt",
  header = F,
  check.names = T,
  stringsAsFactors = F,
  colClasses = "character"
)
colnames(intact_pairs_nodsp_no_dsppmids_mitab_miscore) <- colnames(intact_pairs_dsp_no_dsppmids_mitab)[1:15]

intact_pairs_nodsp_old_mitab_miscore <- fread(
  "./processed_data/intact_pairs_nodsp_old_clustered.txt",
  header = F,
  check.names = T,
  stringsAsFactors = F,
  colClasses = "character"
)
colnames(intact_pairs_nodsp_old_mitab_miscore) <- colnames(intact_pairs_dsp_no_dsppmids_mitab)[1:15]

intact_pairs_nodsp_old_no_dsppmids_mitab_miscore <- fread(
  "./processed_data/intact_pairs_nodsp_old_no_dsppmids_clustered.txt",
  header = F,
  check.names = T,
  stringsAsFactors = F,
  colClasses = "character"
)
colnames(intact_pairs_nodsp_old_no_dsppmids_mitab_miscore) <- colnames(intact_pairs_dsp_no_dsppmids_mitab)[1:15]
```


### Comparison before/after dsp
##### Preformat the data: DSP
```{r miscore_bef_after_dsp}
intact_pairs_dsp_after <- unique(intact_pairs_dsp_mitab_miscore[,.(
  pair_id = apply(data.table(gsub("uniprotkb:","",X.ID.s..interactor.A),
                             gsub("uniprotkb:","",ID.s..interactor.B),
                             stringsAsFactors = F),
                  1,
                  function(a) { z = sort(a)
                  paste0(z[1],"_",z[2]) }),
  miscore_after = as.numeric(gsub(".*miscore:","",Confidence.value.s.)))])

intact_pairs_dsp_before <- unique(intact_pairs_dsp_no_dsppmids_mitab_miscore[,.(
  # p1 = gsub("-.*","",gsub("uniprotkb:","",X.ID.s..interactor.A)),
  # p2 = gsub("-.*","",gsub("uniprotkb:","",ID.s..interactor.B)),
  pair_id = apply(data.table(gsub("uniprotkb:","",X.ID.s..interactor.A),
                             gsub("uniprotkb:","",ID.s..interactor.B),
                             stringsAsFactors = F),
                  1,
                  function(a) { z = sort(a)
                  paste0(z[1],"_",z[2]) }),
  miscore_before = as.numeric(gsub(".*miscore:","",Confidence.value.s.)))])

comp_pairs_dsp_miscore <- unique(merge(
  intact_pairs_dsp_before,
  intact_pairs_dsp_after,
  by = "pair_id",
  all = T
  ))

comp_pairs_dsp_miscore[is.na(miscore_before)]$miscore_before <- 0
comp_pairs_dsp_miscore[,diff:=(miscore_after-miscore_before)]
comp_pairs_dsp_miscore[,set:="dsp"]
```

##### Preformat the data: non-DSP cc-lines
```{r miscore_bef_after_nodsp}
intact_pairs_nodsp_after <- unique(intact_pairs_nodsp_mitab_miscore[,.(
  # p1 = gsub("-.*","",gsub("uniprotkb:","",X.ID.s..interactor.A)),
  # p2 = gsub("-.*","",gsub("uniprotkb:","",ID.s..interactor.B)),
  pair_id = apply(data.table(gsub("uniprotkb:","",X.ID.s..interactor.A),
                             gsub("uniprotkb:","",ID.s..interactor.B),
                             stringsAsFactors = F),
                  1,
                  function(a) { z = sort(a)
                  paste0(z[1],"_",z[2]) }),
  miscore_after = as.numeric(gsub(".*miscore:","",Confidence.value.s.)))])

intact_pairs_nodsp_before <- unique(intact_pairs_nodsp_no_dsppmids_mitab_miscore[,.(
  # p1 = gsub("-.*","",gsub("uniprotkb:","",X.ID.s..interactor.A)),
  # p2 = gsub("-.*","",gsub("uniprotkb:","",ID.s..interactor.B)),
  pair_id = apply(data.table(gsub("uniprotkb:","",X.ID.s..interactor.A),
                             gsub("uniprotkb:","",ID.s..interactor.B),
                             stringsAsFactors = F),
                  1,
                  function(a) { z = sort(a)
                  paste0(z[1],"_",z[2]) }),
  miscore_before = as.numeric(gsub(".*miscore:","",Confidence.value.s.)))])

comp_pairs_nodsp_miscore <- unique(merge(
  intact_pairs_nodsp_before,
  intact_pairs_nodsp_after,
  by = "pair_id",
  all = T
  ))

comp_pairs_nodsp_miscore[is.na(miscore_before)]$miscore_before <- 0
comp_pairs_nodsp_miscore[,diff:=(miscore_after-miscore_before)]
comp_pairs_nodsp_miscore[,set:="control cc-lines"]
```

##### Preformat the data: non-DSP old
```{r miscore_bef_after_nodsp_old}
intact_pairs_nodsp_old_after <- unique(intact_pairs_nodsp_old_mitab_miscore[,.(
  # p1 = gsub("-.*","",gsub("uniprotkb:","",X.ID.s..interactor.A)),
  # p2 = gsub("-.*","",gsub("uniprotkb:","",ID.s..interactor.B)),
  pair_id = apply(data.table(gsub("uniprotkb:","",X.ID.s..interactor.A),
                             gsub("uniprotkb:","",ID.s..interactor.B),
                             stringsAsFactors = F),
                  1,
                  function(a) { z = sort(a)
                  paste0(z[1],"_",z[2]) }),
  miscore_after = as.numeric(gsub(".*miscore:","",Confidence.value.s.)))])

intact_pairs_nodsp_old_before <- unique(intact_pairs_nodsp_old_no_dsppmids_mitab_miscore[,.(
  # p1 = gsub("-.*","",gsub("uniprotkb:","",X.ID.s..interactor.A)),
  # p2 = gsub("-.*","",gsub("uniprotkb:","",ID.s..interactor.B)),
  pair_id = apply(data.table(gsub("uniprotkb:","",X.ID.s..interactor.A),
                             gsub("uniprotkb:","",ID.s..interactor.B),
                             stringsAsFactors = F),
                  1,
                  function(a) { z = sort(a)
                  paste0(z[1],"_",z[2]) }),
  miscore_before = as.numeric(gsub(".*miscore:","",Confidence.value.s.)))])

comp_pairs_nodsp_old_miscore <- unique(merge(
  intact_pairs_nodsp_old_before,
  intact_pairs_nodsp_old_after,
  by = "pair_id",
  all = T
  ))

comp_pairs_nodsp_old_miscore[is.na(miscore_before)]$miscore_before <- 0
comp_pairs_nodsp_old_miscore[,diff:=(miscore_after-miscore_before)]
comp_pairs_nodsp_old_miscore[,set:="control"]
```

#### Explore differences statistics
```{r diff_stats}
comp_pairs_miscore <- unique(rbind(
  comp_pairs_dsp_miscore,
  comp_pairs_nodsp_miscore,
  comp_pairs_nodsp_old_miscore
))

# Wilcoxon test to check if MIscore gain is greater in DSP rather than control or viceversa 
wilcox.test.altg = function(x, y) wilcox.test(x, y, alternative = 'g')
wilcox.test.altl = function(x, y) wilcox.test(x, y, alternative = 'l')
wilcox.test.altg(comp_pairs_miscore[diff!=0 & set == "dsp"]$diff,comp_pairs_miscore[diff!=0 & set == "control"]$diff)
wilcox.test.altl(comp_pairs_miscore[diff!=0 & set == "dsp"]$diff,comp_pairs_miscore[diff!=0 & set == "control"]$diff)
wilcox.test.altg(comp_pairs_miscore[diff!=0 & set == "dsp"]$diff,comp_pairs_miscore[diff!=0 & set == "control cc-lines"]$diff)
wilcox.test.altl(comp_pairs_miscore[diff!=0 & set == "dsp"]$diff,comp_pairs_miscore[diff!=0 & set == "control cc-lines"]$diff)

# Summary statistics

## Old control vs DSP
new_pairs <- data.table(
  dsp = c(nrow(comp_pairs_miscore[set=="dsp" & miscore_before==0]),nrow(comp_pairs_miscore[set=="dsp" & miscore_before!=0])),
  control = c(nrow(comp_pairs_miscore[set=="control" & miscore_before==0]),nrow(comp_pairs_miscore[set=="control" & miscore_before!=0]))
)
new_pairs_matrix <- as.matrix(new_pairs)
fisher.test(new_pairs_matrix,
            alternative="two.sided")

unaff_pairs <- data.table(
  dsp = c(nrow(comp_pairs_miscore[set=="dsp" & diff==0]),nrow(comp_pairs_miscore[set=="dsp" & diff!=0])),
  control = c(nrow(comp_pairs_miscore[set=="control" & diff==0]),nrow(comp_pairs_miscore[set=="control" & diff!=0]))
)

unaff_pairs_matrix <- as.matrix(unaff_pairs)
fisher.test(unaff_pairs_matrix,
            alternative="two.sided")

## CC-lines control vs DSP
new_pairs_cc <- data.table(
  dsp = c(nrow(comp_pairs_miscore[set=="dsp" & miscore_before==0]),nrow(comp_pairs_miscore[set=="dsp" & miscore_before!=0])),
  control_cc = c(nrow(comp_pairs_miscore[set=="control cc-lines" & miscore_before==0]),nrow(comp_pairs_miscore[set=="control cc-lines" & miscore_before!=0]))
)
new_pairs_cc_matrix <- as.matrix(new_pairs_cc)
fisher.test(new_pairs_cc_matrix,
            alternative="two.sided")

unaff_pairs_cc <- data.table(
  dsp = c(nrow(comp_pairs_miscore[set=="dsp" & diff==0]),nrow(comp_pairs_miscore[set=="dsp" & diff!=0])),
  control_cc = c(nrow(comp_pairs_miscore[set=="control cc-lines" & diff==0]),nrow(comp_pairs_miscore[set=="control cc-lines" & diff!=0]))
)

unaff_pairs_cc_matrix <- as.matrix(unaff_pairs_cc)
fisher.test(unaff_pairs_cc_matrix,
            alternative="two.sided")

## CC-lines control vs old control
new_pairs_cc_ctrl <- data.table(
  control = c(nrow(comp_pairs_miscore[set=="control" & miscore_before==0]),nrow(comp_pairs_miscore[set=="control" & miscore_before!=0])),
  control_cc = c(nrow(comp_pairs_miscore[set=="control cc-lines" & miscore_before==0]),nrow(comp_pairs_miscore[set=="control cc-lines" & miscore_before!=0]))
)
new_pairs_cc_ctrl_matrix <- as.matrix(new_pairs_cc_ctrl)
fisher.test(new_pairs_cc_ctrl_matrix,
            alternative="two.sided")

unaff_pairs_cc_ctrl <- data.table(
  control = c(nrow(comp_pairs_miscore[set=="control" & diff==0]),nrow(comp_pairs_miscore[set=="control" & diff!=0])),
  control_cc = c(nrow(comp_pairs_miscore[set=="control cc-lines" & diff==0]),nrow(comp_pairs_miscore[set=="control cc-lines" & diff!=0]))
)

unaff_pairs_cc_ctrl_matrix <- as.matrix(unaff_pairs_cc_ctrl)
fisher.test(unaff_pairs_cc_ctrl_matrix,
            alternative="two.sided")
```
I find no significant change in the amount of new pairs discovered through DSP in comparison with the cc-lines control, but both DSP and cc-lines control show a significant difference with the old control.  

#### Plot differences
```{r diff_plot}
g <- ggplot(comp_pairs_miscore[diff!=0], aes(x=set,y=diff)) +
  geom_violin() +
  geom_jitter(alpha=0.2) +
  ylab('MIscore net gain') + 
  xlab('') +
  geom_signif(comparisons = list(c("control","dsp"),
                                 c("control cc-lines","dsp"),
                                 c("control","control cc-lines")),
  # geom_signif(comparisons = c("dsp","control"),
              step_increase=0.07,
              test = wilcox.test.altg,
              tip_length = 0.015,
              textsize = 5) +
  theme(plot.title = element_text(),
        legend.position = "none",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(),
        axis.text.x = element_text(size=16),
        axis.text.y = element_text(size=16),
        axis.title = element_text(size=18))
g
```



## Correlate interest scores with MIscore progression
If interest score really helps targeting lowly annotated pairs, it should correlate with bigger differences in MIscore progression. This comparison is only possible for DSP-curated papers, of course.  
```{r interest_miscore}
pairs_int_comp <- unique(merge(
  eval_dsp_pairs_formatted,
  comp_pairs_dsp_miscore,
  by.x = "prot_pairs",
  by.y = "pair_id",
  all = F
))

fwrite(
  pairs_int_comp,
  "./processed_data/pairs_int_comp.txt",
  col.names = T,
  row.names = F,
  quote = F,
  sep = "\t"
)
```

### Plots interest scores vs MIscore
```{r plot_intvsmiscore_1}
g1 <- ggplot(pairs_int_comp,aes(x=as.numeric(DS.interest.up),y=diff)) +
  geom_point(alpha = 0.2) + 
  geom_smooth(method='lm') +
  ylab('MIscore net gain') + 
  xlab('DSP interest score') +
  theme(plot.title = element_text(),
        legend.position = "none",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(),
        axis.text.x = element_text(size=16),
        axis.text.y = element_text(size=16),
        axis.title = element_text(size=18))
g1
```

```{r plot_intvsmiscore_2}
g2 <- ggplot(pairs_int_comp,aes(x=as.numeric(DS.part_interest.up),y=diff)) +
  geom_point(alpha = 0.2) + 
  geom_smooth(method='lm') +
  ylab('MIscore net gain') + 
  xlab('DSP interest score, partial') +
  theme(plot.title = element_text(),
        legend.position = "none",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(),
        axis.text.x = element_text(size=16),
        axis.text.y = element_text(size=16),
        axis.title = element_text(size=18))
g2
```

```{r plot_intvsmiscore_3}
g3 <- ggplot(pairs_int_comp,aes(x=as.numeric(DS.score.pred),y=diff)) +
  geom_point(alpha = 0.2) + 
  geom_smooth(method='lm') +
  ylab('MIscore net gain') + 
  xlab('DSP interest score, with predictions') +
  theme(plot.title = element_text(),
        legend.position = "none",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(),
        axis.text.x = element_text(size=16),
        axis.text.y = element_text(size=16),
        axis.title = element_text(size=18))
g3
```

Any version of the interest score is all over the place, with poor correlation with MIscore gains. The correlation seems to be actually negative, if anything. 

## Normalization problem evaluation
Let's check how much of a problem we have with identification of interacting pairs in text mining and predicted datasets. 
```{r normeval}
curated_dsp_sel <-unique(eval_dsp[grepl("yes|Yes",curatable) & pmid %in% publs_curated_dsp$pmid,.(
  pair_id= prot_pairs,
  pmid,
  text_mined = ifelse(
    tm_epmc_pair == "yes" | EVEX_pair == "yes" | STRING_textmining_pair == "yes",
    "yes",
    ifelse(
      tm_epmc_pair == "part" | EVEX_pair == "part" | STRING_textmining_pair == "part",
      "part",
      "no"
    )
  ),
  pw_curated = ifelse(
    reactome_pair == "yes" | OmniPath_interactions_pair == "yes" | OmniPath_ptm_pair == "yes" | STRING_pi_pair == "yes",
    "yes",
    ifelse(
      reactome_pair == "part" | OmniPath_interactions_pair == "part" | OmniPath_ptm_pair == "part" | STRING_pi_pair == "part",
      "part",
      "no"
    )
  ),
  predicted = ifelse(
    iid_pred == "yes",
    "yes",
    ifelse(
       iid_pred == "part",
      "part",
      "no"
    )
  )
  )])

curated_dsp_long <- cSplit(
  curated_dsp_sel,
  splitCols = "pair_id",
  sep = "|",
  direction = "long",
  drop = F
)

intact_pairs_dsp_sel <- intact_human[pmid %in% publs_curated_dsp$pmid,.(
  pmid,
  pair_id,
  imex_curated = "yes"
)]

predicted_vs_curated_dsp_pairs <- unique(merge(
  curated_dsp_long,
  intact_pairs_dsp_sel,
  by = c("pair_id","pmid"),
  all = T
))

predicted_vs_curated_dsp_pairs_annot <- predicted_vs_curated_dsp_pairs[,.(
  pair_id,
  pmid,
  match = ifelse(
    imex_curated == "yes" & is.na(text_mined),
    "not DSP predicted",
    ifelse(
      imex_curated == "yes" & (text_mined == "yes" | pw_curated == "yes" | predicted == "yes"),
      "full match",
      ifelse(
        imex_curated == "yes" & (text_mined == "part" | pw_curated == "part" | predicted == "part"),
        "partial match",
        "false identification"
      )
    )
  ),
  imex_curated,
  text_mined,
  pw_curated,
  predicted
)]

table(predicted_vs_curated_dsp_pairs_annot$match,useNA="ifany")

true_positives_overview <- melt(
  predicted_vs_curated_dsp_pairs_annot[match=="full match"],
  id.vars = c("pmid","pair_id"),
  measure.vars = c("text_mined","pw_curated","predicted"))

table(true_positives_overview$variable,
      true_positives_overview$value,
      useNA = "ifany")

part_positives_overview <- melt(
  predicted_vs_curated_dsp_pairs_annot[match=="partial match"],
  id.vars = c("pmid","pair_id"),
  measure.vars = c("text_mined","pw_curated","predicted"))

table(part_positives_overview$variable,
      part_positives_overview$value,
      useNA = "ifany")

false_positives_overview <- melt(
  predicted_vs_curated_dsp_pairs_annot[is.na(match)],
  id.vars = c("pmid","pair_id"),
  measure.vars = c("text_mined","pw_curated","predicted"))

table(false_positives_overview$variable,
      false_positives_overview$value,
      useNA = "ifany")

table(predicted_vs_curated_dsp_pairs_annot$text_mined,useNA = "ifany")
table(predicted_vs_curated_dsp_pairs_annot$pw_curated,useNA = "ifany")
table(predicted_vs_curated_dsp_pairs_annot$predicted,useNA = "ifany")
```

### Evaluation of mutation data
This is for an unrelated issue: how many curated DSP publications ended up containing mutation data?
```{r muts}
dsp_muts_ints <- intact_dsp[grepl("mutation",Feature.s..interactor.A) | grepl("mutation",Feature.s..interactor.B)]
dsp_muts_publs <- unique(dsp_muts_ints[,.(Publication.Identifier.s.)])
```
There are `r nrow(dsp_muts_ints)` binaries with reported mutations curated, in `r nrow(dsp_muts_publs)` publications. 
**********
<!-- ### Debugging: Negative MIscore progression -->
<!-- Why do I have values with negative MIscore progression? It does not make any sense. I isolate those and run a test.  -->
<!-- ```{r debugmiscore} -->
<!-- test_after <- intact_human[pair_id %in% comp_pairs_dsp_miscore[diff < 0]$pair_id] -->
<!-- test_before <- intact_human[pair_id %in% comp_pairs_dsp_miscore[diff < 0]$pair_id & !(pmid %in% eval_dsp_pmids$pmid)] -->

<!-- test_after_mitab <- test_after[,`:=`(pmid = NULL, p1 = NULL, p2 = NULL, pair_id = NULL)] -->
<!-- fwrite( -->
<!--   test_after_mitab, -->
<!--   "./processed_data/test_after_mitab.txt", -->
<!--   col.names = T, -->
<!--   row.names = F, -->
<!--   quote = F, -->
<!--   sep = "\t" -->
<!-- ) -->

<!-- test_before_mitab <- test_before[,`:=`(pmid = NULL, p1 = NULL, p2 = NULL, pair_id = NULL)] -->
<!-- fwrite( -->
<!--   test_before_mitab, -->
<!--   "./processed_data/test_before_mitab.txt", -->
<!--   col.names = T, -->
<!--   row.names = F, -->
<!--   quote = F, -->
<!--   sep = "\t" -->
<!-- ) -->

<!-- test_before_mitab_miscore <- fread( -->
<!--   "./processed_data/test_before_clustered.txt", -->
<!--   header = F, -->
<!--   check.names = T, -->
<!--   stringsAsFactors = F, -->
<!--   colClasses = "character" -->
<!-- ) -->
<!-- colnames(test_before_mitab_miscore) <- colnames(intact_pairs_dsp_no_dsppmids_mitab)[1:15] -->

<!-- test_after_mitab_miscore <- fread( -->
<!--   "./processed_data/test_after_clustered.txt", -->
<!--   header = F, -->
<!--   check.names = T, -->
<!--   stringsAsFactors = F, -->
<!--   colClasses = "character" -->
<!-- ) -->
<!-- colnames(test_after_mitab_miscore) <- colnames(intact_pairs_dsp_no_dsppmids_mitab)[1:15] -->

<!-- test_pairs_before <- unique(test_before_mitab_miscore[,.( -->
<!--   # p1 = gsub("-.*","",gsub("uniprotkb:","",X.ID.s..interactor.A)), -->
<!--   # p2 = gsub("-.*","",gsub("uniprotkb:","",ID.s..interactor.B)), -->
<!--   pair_id = apply(data.table(gsub("uniprotkb:","",X.ID.s..interactor.A), -->
<!--                              gsub("uniprotkb:","",ID.s..interactor.B), -->
<!--                              stringsAsFactors = F), -->
<!--                   1, -->
<!--                   function(a) { z = sort(a) -->
<!--                   paste0(z[1],"_",z[2]) }), -->
<!--   miscore_before = as.numeric(gsub(".*miscore:","",Confidence.value.s.)))]) -->

<!-- test_pairs_after <- unique(test_after_mitab_miscore[,.( -->
<!--   # p1 = gsub("-.*","",gsub("uniprotkb:","",X.ID.s..interactor.A)), -->
<!--   # p2 = gsub("-.*","",gsub("uniprotkb:","",ID.s..interactor.B)), -->
<!--   pair_id = apply(data.table(gsub("uniprotkb:","",X.ID.s..interactor.A), -->
<!--                              gsub("uniprotkb:","",ID.s..interactor.B), -->
<!--                              stringsAsFactors = F), -->
<!--                   1, -->
<!--                   function(a) { z = sort(a) -->
<!--                   paste0(z[1],"_",z[2]) }), -->
<!--   miscore_after = as.numeric(gsub(".*miscore:","",Confidence.value.s.)))]) -->

<!-- comp_pairs_test_miscore <- unique(merge( -->
<!--   test_pairs_before, -->
<!--   test_pairs_after, -->
<!--   by = "pair_id", -->
<!--   all = T -->
<!--   )) -->

<!-- comp_pairs_test_miscore[is.na(miscore_before)]$miscore_before <- 0 -->
<!-- comp_pairs_test_miscore[,diff:=(miscore_after-miscore_before)] -->
<!-- ``` -->

<!-- Problem is that I need to clean up isoforms! -->

