---
output: 
  html_document: 
    keep_md: yes
---
EVEX dataset generator
========================================================

#### Load EVEX data

```{r echo=FALSE}
EVEX_date <- date()
suppressPackageStartupMessages({
        library(data.table)
        library(VennDiagram)
        library(downloader)
        library(R.utils)
        library(XML)
        library(UniProt.ws)
        })
```

I download the latest version of the EVEX data from EVEX website: relations table and articles table (which have common General event ID). Current file was downloaded on `r EVEX_date`. 

Format rescription:

```{r message=FALSE, warning=FALSE}
README = readLines("http://evexdb.org/download/network-format/README-network.txt")
README[19:35]
README[39:45]
```

```{r scrape_and_download_EVEX, message=FALSE, warning=FALSE}
url_EVEX = "http://evexdb.org/download/network-format/"
directories = htmlTreeParse(url_EVEX, useInternalNodes = T)
directories = xpathSApply(directories, "//a", xmlValue)
directories = grep("^[[:alpha:]]+/", directories , value = T)
filepaths = character()
files = character()
for(directory in directories) {
        files_temp = htmlTreeParse(paste0(url_EVEX, directory), useInternalNodes = T)
        files_temp = xpathSApply(files_temp, "//a", xmlValue)
        files_temp = grep("\\.tar\\.gz$", files_temp , value = T)
        files = c(files, files_temp)
        filepaths_temp = paste0(url_EVEX, directory,files_temp)
        filepaths = c(filepaths, filepaths_temp)
}
files = paste0("./source_files/", files)
for(i in 1:length(filepaths)){
       if(!file.exists(files[i])) download(filepaths[i],files[i])
}

```

```{r message=FALSE, warning=FALSE}
homo_file = grep("Homo", files, value =T)
gunzip(homo_file, remove = F)
untar(substr(homo_file, 1, nchar(homo_file)-3), exdir = "./source_files/Homo/")
unlink(substr(homo_file, 1, nchar(homo_file)-3))
Homo_sapiens_EVEX_relations = fread("./source_files/Homo/EVEX_relations_9606.tab", colClasses = "character")
Homo_sapiens_EVEX_articles = fread("./source_files/Homo/EVEX_articles_9606.tab", colClasses = "character")
```

EVEX is a text-mining resource which aims to identify interactions of different types from the literature as well as segment those interactions by type and measure the confidence of those in interactions being really described in the articles (not an artifact of text-mining).   
An interaction is identified by a pair of genes (in the network format, there is another data format in EVEX which is not relevant here), segmented by type and polarity, and given a confidence score.

Below you can see how many interactions belong to each type. 

```{r types_of_interactions}
Homo_sapiens_EVEX_relations[,table(coarse_type)]
Homo_sapiens_EVEX_relations[,table(refined_type,refined_polarity)]
```

The total number of articles analysed: `r length(unique(Homo_sapiens_EVEX_articles$article_id))`  

```{r unique_pairs, message=FALSE, warning=FALSE, results='hide'}
pairs_entrezgene_id = Homo_sapiens_EVEX_relations[, as.character(unique(c(source_entrezgene_id, target_entrezgene_id)))]
# connecting to Uniprot
 uniprot = UniProt.ws(taxId=9606)
# getting names of the keys and columns from Uniprot
# keytypes(uniprot)
# mapping IDs
 entrezgene2uniprot_temp <- as.data.table(select(uniprot, 
               keys = pairs_entrezgene_id, 
               columns = c("ENTREZ_GENE", "UNIPROTKB"),
               keytype = "ENTREZ_GENE"))
 entrezgene2uniprot = copy(entrezgene2uniprot_temp[complete.cases(entrezgene2uniprot_temp),])
# what is the fraction of matches
# mean(complete.cases(entrezgene2uniprot_temp)) # 0.9517783
# what is the fraction of one to many mappings to UNIPROTKB?
# sum(table(entrezgene2uniprot_comp_temp[, .N , by = ENTREZ_GENE]$N)[-1])/sum(table(entrezgene2uniprot_comp_temp[, .N , by = ENTREZ_GENE]$N)) # 0.5032162

# connecting to ENSEMBL
# ensembl=useMart("ensembl")
# ensembl = useDataset("hsapiens_gene_ensembl",mart=ensembl)
# getting names of the keys and columns from ENSEMBL
# attributes = listAttributes(ensembl)
# entrezgene_uniprot_ENSEMBL = data.table()
# for(i in seq(1,length(pairs_entrezgene_id),500)){
# entrezgene_temp = as.data.table(getBM(attributes = c('entrezgene', 'uniprot_swissprot',"uniprot_sptrembl"), 
#               values = pairs_entrezgene_id[i:(i+499)], filters = 'entrezgene',
#               mart = ensembl))
# entrezgene_uniprot_ENSEMBL = rbind(entrezgene_uniprot_ENSEMBL, entrezgene_temp)
# }
# entrezgene_uniprot_ENSEMBL = unique(entrezgene_uniprot_ENSEMBL)

# what is the fraction of matches?
# 1-mean(entrezgene_uniprot_ENSEMBL$uniprot_swissprot == "" & entrezgene_uniprot_ENSEMBL$uniprot_sptrembl == "") # 0.9951115

# what is the fraction of one to many mappings to SwissProt+trEMBL?
# sum(table(entrezgene_uniprot_ENSEMBL[, .N , by = entrezgene]$N)[-1])/sum(table(entrezgene_uniprot_ENSEMBL[, .N , by = entrezgene]$N)) # 0.7543455

# what is the fraction of one to many mappings to SwissProt?
# entrezgene_uniprot_ENSEMBL_z = unique(entrezgene_uniprot_ENSEMBL[, .(entrezgene, uniprot_swissprot)])
# sum(table(entrezgene_uniprot_ENSEMBL_z[, .N , by = entrezgene]$N)[-1])/sum(table(entrezgene_uniprot_ENSEMBL_z[, .N , by = entrezgene]$N)) # 0.7310322

Homo_sapiens_EVEX_relations = unique(Homo_sapiens_EVEX_relations)
entrezgene2uniprot = unique(entrezgene2uniprot)
Homo_sapiens_EVEX_relations = merge(x = Homo_sapiens_EVEX_relations, y = entrezgene2uniprot, by.x = "source_entrezgene_id", by.y = "ENTREZ_GENE", all.x = T, allow.cartesian=TRUE)
setnames(Homo_sapiens_EVEX_relations, colnames(Homo_sapiens_EVEX_relations)[ncol(Homo_sapiens_EVEX_relations)], "source_uniprotkb_id")
Homo_sapiens_EVEX_relations = merge(x = Homo_sapiens_EVEX_relations, y = entrezgene2uniprot, by.x = "target_entrezgene_id", by.y = "ENTREZ_GENE", all.x = T, allow.cartesian=TRUE)
setnames(Homo_sapiens_EVEX_relations, colnames(Homo_sapiens_EVEX_relations)[ncol(Homo_sapiens_EVEX_relations)], "target_uniprotkb_id")

# filter out interactions with unmapped source or target id
Homo_sapiens_EVEX_relations = Homo_sapiens_EVEX_relations[!(is.na(source_uniprotkb_id) | is.na(target_uniprotkb_id)),]
# filter for interaction types
interaction_types = c("Binding", "Catalysis of acetylation", "Catalysis of DNA methylation", "Catalysis of glycosylation","Catalysis of hydroxylation", "Catalysis of methylation", "Catalysis of phosphorylation", "Catalysis of ubiquitination")
EVEX_pairs = Homo_sapiens_EVEX_relations[refined_type %in% interaction_types, 
                                                          .(general_event_id, ida_clean = source_uniprotkb_id,idb_clean = target_uniprotkb_id, interaction_type = refined_type, interaction_confidence = confidence, taxon = "9606", EVEX = 1)]
# generating interacting pairs
EVEX_pairs[, pair_id := apply(data.table(ida_clean,idb_clean,stringsAsFactors = F), 1,
                                               function(a) { z = sort(a)
                                               paste0(z[1],"_",z[2]) })]

# Mapping between general_event_id and publicationID
EVEX_pairs[, pmids_mixed := Homo_sapiens_EVEX_articles$article_id[match(general_event_id, Homo_sapiens_EVEX_articles$general_event_id)]]
merge(x = EVEX_pairs, y = Homo_sapiens_EVEX_articles, by.x = "general_event_id", by.y = "general_event_id", all.x = T)
EVEX_pairs[, pmids_mixed := gsub("PMID: ", "",pmids_mixed)]
EVEX_pairs[, pmids_mixed := gsub("PMCID: ", "",pmids_mixed)]
# Downloading PMCID to PMID conversion table
url = "ftp://ftp.ncbi.nlm.nih.gov/pub/pmc/PMC-ids.csv.gz"
filename = "./source_files/PMC-ids.csv.gz"
if(!file.exists(filename)) download(url,filename); 
gunzip(filename, remove = F)
PMCID_2_PMID = fread(substr(filename,1, nchar(filename)-3), colClasses = "character")
unlink(substr(filename,1, nchar(filename)-3))
EVEX_pairs[, pmids := PMCID_2_PMID$PMID[match(pmids_mixed, PMCID_2_PMID$PMCID)]]
EVEX_pairs[is.na(pmids), pmids := pmids_mixed]

# saving detailed EVEX table with standard columns
EVEX_pairs = EVEX_pairs[, .(pair_id_clean = pair_id, ida_clean, idb_clean, pubid = pmids, taxon, interaction_type, interaction_confidence, EVEX)]
fwrite(x = unique(EVEX_pairs), 
       file = "./results/pairs_pmids_EVEX_detailed.txt", sep = "\t")
N_EVEX = length(EVEX_pairs[,unique(pair_id_clean)])

# saving minimal EVEX table with standard columns
EVEX_pairs_s = EVEX_pairs[, .(pair_id_clean, ida_clean, idb_clean, pubid, taxon, EVEX, evex_score = interaction_confidence)]
fwrite(x = unique(EVEX_pairs_s), 
       file = "./results/pairs_pmids_EVEX_minimal.txt", sep = "\t")
```

The total number of interacting pairs extracted from EVEX:`r N_EVEX`

I create a list of PMIDs that have been mined by them.

```{r pmids}
EVEX_pmids <- data.frame(unique(EVEX_pairs$pubid))
write.table(EVEX_pmids, "./results/EVEX_pmids.txt", quote=F, sep ="\t", row.names = F, col.names = T)
```

`r nrow(EVEX_pmids)` publications are were mined to get interactions in the EVEX database. 

#### Compare EVEX interactions and publications to IMEx 

I calculate how many interactions in EVEX match to IMEx.

```{r biogrid_vs_imex, message=FALSE, warning=FALSE, fig.width=4, fig.height=2}
imex = fread("https://raw.githubusercontent.com/pporrasebi/darkspaceproject/master/IMEx/results/imex_full.txt", header = T, sep = "\t", colClasses = "character")
imex_human = imex[taxid_a == "9606" | taxid_b == "9606",]
N_imex = length(imex_human[,unique(pair_id_clean)])
N_EVEX = length(EVEX_pairs[,unique(pair_id_clean)])
N_overlap = sum(!is.na(match(EVEX_pairs[,unique(pair_id_clean)], imex_human[,unique(pair_id_clean)])))

venn.d = draw.pairwise.venn(area1 = N_imex, area2 = N_EVEX, cross.area = N_overlap, category = c("IMEx", "EVEX"), 
                          lty = rep("blank", 2), 
                          fill = c("blue", "red"), 
                          alpha = rep(0.5, 2), cat.pos = c(0, 135), 
                          cat.dist = rep(0.035, 2), 
                          cat.cex = c(1,1), scaled = TRUE, euler.d = TRUE,  margin = 0.05,
                          direct.area = TRUE,
                          cex = 1)

```

I calculate how many publications in EVEX match to IMEx.

```{r biogrid_vs_imex_pub, message=FALSE, warning=FALSE, fig.width=4, fig.height=2}
N_pubid_imex = length(imex_human[,unique(pubid)])
N_pubid_EVEX = length(EVEX_pairs[,unique(pubid)])
N_pubid_overlap = sum(!is.na(match(EVEX_pairs[,unique(pubid)], imex_human[,unique(pubid)])))

venn.d = draw.pairwise.venn(area1 = N_pubid_imex, area2 = N_pubid_EVEX, cross.area = N_pubid_overlap, category = c("IMEx", "EVEX"), 
                          lty = rep("blank", 2), 
                          fill = c("blue", "red"), 
                          alpha = rep(0.5, 2), cat.pos = c(0, 135), 
                          cat.dist = rep(0.035, 2), 
                          cat.cex = c(1,1), scaled = TRUE, euler.d = TRUE,  margin = 0.05,
                          direct.area = TRUE,
                          cex = 1)

```

I calculate how many interactions published in specific articles (the same interaction can have evidence from different publications) in EVEX match to IMEx.

```{r biogrid_vs_imex_pub_inter, message=FALSE, warning=FALSE, fig.width=4, fig.height=2}
N_pub_int_imex = length(imex_human[,unique(paste0(pubid,"_",pair_id_clean))])
N_pub_int_EVEX = length(EVEX_pairs[,unique(paste0(pubid,"_",pair_id_clean))])
N_pub_int_overlap = sum(!is.na(match(EVEX_pairs[,unique(paste0(pubid,"_",pair_id_clean))], imex_human[,unique(paste0(pubid,"_",pair_id_clean))])))

venn.d = draw.pairwise.venn(area1 = N_pub_int_imex, area2 = N_pub_int_EVEX, cross.area = N_pub_int_overlap, category = c("IMEx", "EVEX"), 
                          lty = rep("blank", 2), 
                          fill = c("blue", "red"), 
                          alpha = rep(0.5, 2), cat.pos = c(0, 135), 
                          cat.dist = rep(0.035, 2), 
                          cat.cex = c(1,1), scaled = TRUE, euler.d = TRUE,  margin = 0.05,
                          direct.area = TRUE,
                          cex = 1)

```