---
output: 
  html_document: 
    keep_md: yes
---
BioGRID dataset generator
========================================================

#### Load BioGRID data

```{r echo=FALSE}
BioGRID_mentha_date <- date()
suppressPackageStartupMessages({
        library(data.table)
        library(VennDiagram)
        library(downloader)
        library(R.utils)
        })
```

I download the latest version of the BioGRID data from mentha website. I have tried using PSICQUIC, however, downloading whole mentha using PSICQUIC takes too much time. Now, however, the link should be updated manually to get the latest mentha release. Current file was downloaded on `r BioGRID_mentha_date`. 

```{r message=FALSE, warning=FALSE}
if(!file.exists("./source_files/2017-02-27_MITAB-2.5.zip")){
        mentha_url = "http://mentha.uniroma2.it/dumps/2017-02-27_MITAB-2.5.zip"
        # httr::GET(http://mentha.uniroma2.it/doDownload.php?file=2017-02-27_MITAB-2.5.zip) - this is how I got the direct link
        download(mentha_url, destfile = "./source_files/2017-02-27_MITAB-2.5.zip")
}
unzip("./source_files/2017-02-27_MITAB-2.5.zip", exdir = "./source_files/")
# databases <- c("mentha")
## Query PSICQUIC for interactions, get MI-TAB-2.5, save
# source("./scripts/query_PSICQUIC_for_interactions.R")
# query_PSICQUIC_for_interactions(SPECIES_ID = "all", SPECIES_NAME = "all", databases= "mentha", date = BioGRID_mentha_date,
#                                          database.name = "./source_files/mentha_mitab25.txt", return_data = F, show_summary = F, MITAB = "tab25")
```

Filtering BioGRID only interactions and cleaning up the BioGRID file.

```{r message=FALSE, warning=FALSE}
mentha = fread("./source_files/2017-02-27_MITAB-2.5", header = F, sep = "\t", colClasses = "character")
biogrid_from_mentha = mentha[V13 == "psi-mi:\"MI:0463\"(biogrid)",]
fwrite(x = biogrid_from_mentha, file = "./processed_files/biogrid_mitab25.txt",sep = "\t")
system("perl ./scripts/MITAB25extractor_v12.pl ./processed_files/biogrid_mitab25.txt ./processed_files/biogrid_pairs.txt")
unlink(c("./source_files/2017-02-27_MITAB-2.5","./processed_files/biogrid_mitab25.txt"))
```

Saving a table of interacting pairs, publication IDs and BioGRID tag. Selecting human protein-human protein only interactions.

```{r pairs, message=FALSE, warning=FALSE}
biogrid_from_mentha = fread("./processed_files/biogrid_pairs.txt", header = T, sep = "\t", colClasses = "character")
biogrid_from_mentha_human = biogrid_from_mentha[taxid_a == "9606" | taxid_b == "9606",]
fwrite(x = unique(biogrid_from_mentha_human[, .(pair_id_clean, pubid, biogrid = rep(1, .N))]), 
       file = "./results/pairs_pmids_biogrid.txt", sep = "\t")
N_biogrid = length(biogrid_from_mentha_human[,unique(pair_id_clean)])
```

The BioGRID dataset contains `r N_biogrid` human interacting pairs. 

Files in the /processed_files/ contain interactions for all species.
Files in the /results/ contain interactions for human only.

Creating a list of PMIDs that have been curated into human-human, human-other_species, human-compound interaction

```{r pmids}
biogrid_pmids <- data.frame(unique(biogrid_from_mentha_human$pubid))

write.table(biogrid_pmids, "./results/biogrid_pmids.txt", quote=F, sep ="\t", row.names = F, col.names = T)
```

`r nrow(biogrid_pmids)` publications (human) are curated into BioGRID database. 

#### Compare human BioGRID interactions and publications to IMEx 

I calculate how many interactions in BioGRID match to IMEx.

```{r biogrid_vs_imex, message=FALSE, warning=FALSE, fig.width=4, fig.height=2}
imex = fread("https://raw.githubusercontent.com/pporrasebi/darkspaceproject/master/IMEx/results/imex_full.txt", header = T, sep = "\t", colClasses = "character")
imex_human = imex[taxid_a == "9606" | taxid_b == "9606",]
N_imex = length(imex_human[, unique(pair_id_clean)])
N_biogrid = length(biogrid_from_mentha_human[,unique(pair_id_clean)])
N_overlap = sum(!is.na(match(biogrid_from_mentha_human[,unique(pair_id_clean)], imex_human[, unique(pair_id_clean)])))

venn.d = draw.pairwise.venn(area1 = N_imex, area2 = N_biogrid, cross.area = N_overlap, category = c("IMEx", "BioGRID"), 
                          lty = rep("blank", 2), 
                          fill = c("blue", "red"), 
                          alpha = rep(0.5, 2), cat.pos = c(0, 135), 
                          cat.dist = rep(0.035, 2), 
                          cat.cex = c(1,1), scaled = TRUE, euler.d = TRUE,  margin = 0.05,
                          direct.area = TRUE,
                          cex = 1)

```

I calculate how many publications in BioGRID match to IMEx.

```{r biogrid_vs_imex_pub, message=FALSE, warning=FALSE, fig.width=4, fig.height=2}
N_pubid_imex = length(imex_human[, unique(pubid)])
N_pubid_biogrid = length(biogrid_from_mentha_human[,unique(pubid)])
N_pubid_overlap = sum(!is.na(match(biogrid_from_mentha_human[,unique(pubid)], imex_human[, unique(pubid)])))

venn.d = draw.pairwise.venn(area1 = N_pubid_imex, area2 = N_pubid_biogrid, cross.area = N_pubid_overlap, category = c("IMEx", "BioGRID"), 
                          lty = rep("blank", 2), 
                          fill = c("blue", "red"), 
                          alpha = rep(0.5, 2), cat.pos = c(0, 135), 
                          cat.dist = rep(0.035, 2), 
                          cat.cex = c(1,1), scaled = TRUE, euler.d = TRUE,  margin = 0.05,
                          direct.area = TRUE,
                          cex = 1)

```

I calculate how many interactions published in specific articles (the same interaction can come from different publications) in BioGRID match to IMEx.

```{r biogrid_vs_imex_pub_inter, message=FALSE, warning=FALSE, fig.width=4, fig.height=2}
N_pub_int_imex = length(imex_human[, unique(paste0(pubid,"_",pair_id_clean))])
N_pub_int_biogrid = length(biogrid_from_mentha_human[,unique(paste0(pubid,"_",pair_id_clean))])
N_pub_int_overlap = sum(!is.na(match(biogrid_from_mentha_human[,unique(paste0(pubid,"_",pair_id_clean))], imex_human[, unique(paste0(pubid,"_",pair_id_clean))])))

venn.d = draw.pairwise.venn(area1 = N_pub_int_imex, area2 = N_pub_int_biogrid, cross.area = N_pub_int_overlap, category = c("IMEx", "BioGRID"), 
                          lty = rep("blank", 2), 
                          fill = c("blue", "red"), 
                          alpha = rep(0.5, 2), cat.pos = c(0, 135), 
                          cat.dist = rep(0.035, 2), 
                          cat.cex = c(1,1), scaled = TRUE, euler.d = TRUE,  margin = 0.05,
                          direct.area = TRUE,
                          cex = 1)

```
